### 2019-07-25 figures/approach/evaluation/evaluationDiagram.tex
Add evaluation diagram

\begin{figure}[t]
\begin{centering}
    {\includegraphics[scale=0.9]{figures/approach/evaluation/evaluation-diagram.pdf}}
    \caption[Evaluation against DefinitelyTyped Repository]{\textbf{Evaluation of generated declaration files against DefinitelyTyped Repository} - A parser transforms the generated declaration file and the equivalent file in the DefinitelyTyped repository into a JSON file using the TypeScript Compiler API \citep{typescript-compiler-api}. Comparison is then performed on the JSON files, i.e. not on the declaration files.}
    \label{fig:evaluation-diagram}
\end{centering}
\end{figure}