### 2019-07-27 figures/experiments/operators/relational-operators/relationalOperators.tex

\begin{figure}[h]
	\centering

	\begin{tikzpicture}
		\begin{axis}[
			view={0}{90},
			title=Operator ${<}$,
			width=0.45\textwidth,
			colormap/hot,
			xticklabels={number,string,undefined,object,null,boolean,function,array},
			xtick={0,...,7},
			yticklabels={number,string,undefined,object,null,boolean,function,array},
			ytick={7,...,0},
			x tick label style={rotate=90,anchor=east}]
		]
		\addplot3[surf, shader=interp, left] file {figures/experiments/operators/relational-operators/operator_<.dat};
		\end{axis}
	\end{tikzpicture}
	%
	\begin{tikzpicture}
		\begin{axis}[
			view={0}{90},
			width=0.45\textwidth,
			title=Operator ${>}$,
			colormap/hot,
			xticklabels={number,string,undefined,object,null,boolean,function,array},
			xtick={0,...,7},
			yticklabels={number,string,undefined,object,null,boolean,function,array},
			ytick={7,...,0},
			x tick label style={rotate=90,anchor=east}]
		\addplot3[surf, shader=interp, left] file {figures/experiments/operators/relational-operators/operator_>.dat};
		\end{axis}
	\end{tikzpicture}

	\centering
	\begin{tikzpicture}
		\begin{axis}[
			view={0}{90},
			width=0.45\textwidth,
			title=Operator ${<=}$,
			colormap/hot,
			xticklabels={number,string,undefined,object,null,boolean,function,array},
			xtick={0,...,7},
			yticklabels={number,string,undefined,object,null,boolean,function,array},
			ytick={7,...,0},
			x tick label style={rotate=90,anchor=east}]
		\addplot3[surf, shader=interp, left] file {figures/experiments/operators/relational-operators/operator_<=.dat};
		\end{axis}
	\end{tikzpicture}
	%
	\begin{tikzpicture}
		\begin{axis}[
			view={0}{90},
			width=0.45\textwidth,
			title=Operator ${>=}$,
			colormap/hot,
			xticklabels={number,string,undefined,object,null,boolean,function,array},
			xtick={0,...,7},
			yticklabels={number,string,undefined,object,null,boolean,function,array},
			ytick={7,...,0},
			x tick label style={rotate=90,anchor=east}]
		\addplot3[surf, shader=interp, left] file {figures/experiments/operators/relational-operators/operator_>=.dat};
		\end{axis}
	\end{tikzpicture}

	\centering
	\begin{tikzpicture}
		\begin{axis}[
			view={0}{90},
			width=0.45\textwidth,
			title=Operator ${instanceof}$,
			colormap/hot,
			xticklabels={number,string,undefined,object,null,boolean,function,array},
			xtick={0,...,7},
			yticklabels={number,string,undefined,object,null,boolean,function,array},
			ytick={7,...,0},
			x tick label style={rotate=90,anchor=east}]
		\addplot3[surf, shader=interp, left] file {figures/experiments/operators/relational-operators/operator_instanceof.dat};
		\end{axis}
	\end{tikzpicture}
	%
	\begin{tikzpicture}
		\begin{axis}[
			view={0}{90},
			width=0.45\textwidth,
			title=Operator ${in}$,
			colormap/hot,
			xticklabels={number,string,undefined,object,null,boolean,function,array},
			xtick={0,...,7},
			yticklabels={number,string,undefined,object,null,boolean,function,array},
			ytick={7,...,0},
			x tick label style={rotate=90,anchor=east}]
		\addplot3[surf, shader=interp, left] file {figures/experiments/operators/relational-operators/operator_in.dat};
		\end{axis}
	\end{tikzpicture}

	\centering
	\hspace{0.1\textwidth}
	\begin{tikzpicture}
		\begin{axis}[
			hide axis,
			scale only axis,
			height=0pt,
			width=0pt,
			colormap/hot,
			colorbar horizontal,
			point meta min=0,
			point meta max=100,
			colorbar style={
				title=\%,
				width=0.7\textwidth,
				xtick={0, 10, 20, 30, 40, ..., 100}
			}]
			\addplot [draw=none] coordinates {(0,0)};
		\end{axis}
	\end{tikzpicture}
	\caption[Relational operators]{\textbf{Type distribution for relational operators} - Operators ${<}$, ${>}$, ${<=}$, ${>=}$ are mainly used with the \textit{number-number} combination}
\end{figure}