### 2019-07-29 figures/experiments/overallFunnel.tex
Add funnel with number of modules for each stage of the experiment

\definecolor{myellow}{RGB}{228,212,0}
\definecolor{mgreen}{RGB}{5,104,57}

\newcommand\funnel[3]{%
\pgfmathsetmacro\mwid{(0.3+\val*0.06)}
\pgfmathsetmacro\mradius{(\val*0.01 + 1)}
\pgfmathsetmacro\mheight{(\val*0.003 + 0.4)}
\pgfmathsetmacro\marc{\mwid-.4}
    \begin{scope}[%
        shift={(0,#1)}, 
        line width=.05pt, 
        %x=5mm, 
        %scale=1.\xi,
        yshift=\xi*4
        ]
    \draw[black,bottom color=#2, top color=#2] (-\mwid,0) -- (-\mwid+.4,-\mheight) arc (190:350:\marc cm and \mradius mm) -- (\mwid,0);
    \draw[black,fill=#3] (0,0) ellipse (\mwid cm and \mradius mm);
    \path (-\mwid,0) -- (-\mwid+.4,-\mheight) coordinate[midway] (a\xi);
    \end{scope}
}

\begin{figure}[h]
	\hspace*{-0.28\textwidth}
	\centering
	\begin{tikzpicture}
		\foreach \val
				[%
				count=\xi starting from 0, 
				evaluate=\xi as \shadecolor using int(25*\xi),
				evaluate=\xi as \coord using int(\xi-12)
				]
			in {
				4.05,
				9.37,
				24.02,
				37.49,
				71.19,
				82.50,
				100.00
			}{
				\funnel{\coord}{mgreen!\shadecolor !myellow}{mgreen!\shadecolor !myellow}
			}

		\node[left=0.02\textwidth of a0] {Generated Declaration Files};
		\node[right=0.07\textwidth of a0] {\textbf{244}};

		\node[left=0.02\textwidth of a1] {Run-time Information};
		\node[right=0.12\textwidth of a1] {\textbf{565}};

		\node[left=0.02\textwidth of a2] {Working Examples};
		\node[right=0.26\textwidth of a2] {\textbf{1448}};

		\node[left=0.02\textwidth of a3] {Code Examples};
		\node[right=0.38\textwidth of a3] {\textbf{2260}};

		\node[left=0.02\textwidth of a4] {README file};
		\node[right=0.70\textwidth of a4] {\textbf{4292}};

		\node[left=0.02\textwidth of a5] {Github Repository};
		\node[right=0.80\textwidth of a5] {\textbf{4974}};

		\node[align=right,left=0.02\textwidth of a6, text width=0.25\textwidth] {Definitely Typed Modules};
		\node[right=0.97\textwidth of a6] {\textbf{6029}};

	\end{tikzpicture}
	\caption[Number of analyzed modules for each stage of the experiment]{\textbf{Number of analyzed modules for each stage of the experiment} - A TypeScript Declaration File was generated for only 244 modules, out of 6029 modules in the DefinitelyTyped Repository. It was possible to gather valid run-time information for only 25\% of the modules for which a Code Example was extracted.}
\end{figure}