### 2019-08-05 figures/approach/run-time-information-gathering/analysis/get-field.tex

\begin{figure}[h]
	\centering
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.4\textwidth}
			\jscode{code/run-time-information-gathering/simple-examples/get-field.js}
		\end{minipage}
	\end{lrbox}
	\subfloat[Instrumented JS code]{\usebox{\mintedbox}}
	\hfill
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.58\textwidth}
			\jsoncode{code/run-time-information-gathering/simple-examples/output-get-field.json}
		\end{minipage}
	\end{lrbox}
	\subfloat[Run-time information output]{\usebox{\mintedbox}}
	\caption[\textit{getField} interaction example]{\textbf{\textit{getField} interaction example} - The accessed properties of argument \textit{a} with their type are recorded. Several additional fields and entries of the final output JSON file were ommited for simplicity.}
	\label{fig:run-time-information-gathering-get-field}
\end{figure}