### 2019-08-05 figures/approach/run-time-information-gathering/analysis/method-call.tex

\begin{figure}[h]
	\centering
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.4\textwidth}
			\jscode{code/run-time-information-gathering/simple-examples/method-call.js}
		\end{minipage}
	\end{lrbox}
	\subfloat[Instrumented JS code]{\usebox{\mintedbox}}
	\hfill
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.58\textwidth}
			\jsoncode{code/run-time-information-gathering/simple-examples/output-method-call.json}
		\end{minipage}
	\end{lrbox}
	\subfloat[Run-time information output]{\usebox{\mintedbox}}
	\caption[\textit{methodCall} interaction example]{\textbf{\textit{methodCall} interaction example} - The invoked method of argument \textit{a} is recorded. Method information can be accessed through the reference \mintinline{bash}{functionId}, for which, in this example, the \mintinline{bash}{inputValue} interactions are stored. Main entry for \mintinline{bash}{functionId_3} has been ommited. Several additional fields and entries of the final output JSON file were also ommited for simplicity.}
	\label{fig:run-time-information-gathering-method-call}
\end{figure}