### 2019-08-06 figures/approach/run-time-information-gathering/analysis/following-interactions.tex
Add following interactions and traces examples

\begin{figure}[h]
	\centering
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.4\textwidth}
			\jscode{code/run-time-information-gathering/simple-examples/following-interactions.js}
		\end{minipage}
	\end{lrbox}
	\subfloat[Instrumented JS code]{\usebox{\mintedbox}}
	\hfill
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.58\textwidth}
			\jsoncode{code/run-time-information-gathering/simple-examples/output-following-interactions.json}
		\end{minipage}
	\end{lrbox}
	\subfloat[Run-time information output]{\usebox{\mintedbox}}
	\caption[\textit{followingInteractions} interaction example]{\textbf{\textit{followingInteractions} interaction example} - Observations on return values of \mintinline{bash}{getField} and \mintinline{bash}{methodCall} interactions get associated to the leading interaction. Several additional fields and entries of the final output JSON file were ommited for simplicity.}
	\label{fig:run-time-information-gathering-following-interactions}
\end{figure}