### 2019-08-06 figures/approach/run-time-information-gathering/analysis/traces.tex

\begin{figure}[h]
	\centering
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.46\textwidth}
			\jscode{code/run-time-information-gathering/simple-examples/traces.js}
		\end{minipage}
	\end{lrbox}
	\subfloat[Instrumented JS code]{\usebox{\mintedbox}}
	\hfill
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.52\textwidth}
			\jsoncode{code/run-time-information-gathering/simple-examples/output-traces.json}
		\end{minipage}
	\end{lrbox}
	\subfloat[Run-time information output]{\usebox{\mintedbox}}
	\caption[\textit{traces} example]{\textbf{\textit{traces} example} - Interactions on different invocations of a function are stored with different \mintinline{bash}{traces}. Several additional fields and entries of the final output JSON file were ommited for simplicity.}
	\label{fig:run-time-information-gathering-traces}
\end{figure}