### 2019-08-07 figures/approach/run-time-information-gathering/analysis/operator.tex

\begin{figure}[h]
	\centering
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.34\textwidth}
			\jscode{code/run-time-information-gathering/simple-examples/operator.js}
		\end{minipage}
	\end{lrbox}
	\subfloat[Instrumented JS code]{\usebox{\mintedbox}}
	\hfill
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.62\textwidth}
			\jsoncode{code/run-time-information-gathering/simple-examples/output-operator.json}
		\end{minipage}
	\end{lrbox}
	\subfloat[Run-time information output]{\usebox{\mintedbox}}
	\caption[\textit{operator} interaction example]{\textbf{\textit{operator} interaction example} - The interaction records the operator and the type of both operands. The field \mintinline{bash}{operandForInteraction} indicates which operand triggered the interaction. The example shows how that value is \mintinline{bash}{left} for one argument and \mintinline{bash}{right} for the other one. Several additional fields and entries of the final output JSON file were ommited for simplicity.}
	\label{fig:run-time-information-gathering-following-interactions}
\end{figure}