### 2019-08-10 figures/approach/typescript-declaration-files-generation-method/examples/generation-declaration-file-with-nested-interfaces.tex
Add example for nested interfaces

\begin{figure}[h]
	\centering
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.7\textwidth}
			\jscode{code/typescript-declaration-file-generation/example-with-nested-interfaces/printer.js}
		\end{minipage}
	\end{lrbox}
	\subfloat[printer.js]{\usebox{\mintedbox}}

	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.9\textwidth}
			\tscode{code/typescript-declaration-file-generation/example-with-nested-interfaces/printer.d.ts}
		\end{minipage}
	\end{lrbox}
	\subfloat[printer/index.d.ts]{\usebox{\mintedbox}}

	\caption[Nested interfaces]{\textbf{Nested interfaces} - Sub-interfaces for fields \mintinline{bash}{name} and \mintinline{bash}{address} are declared as separated interfaces and then associated to the main \mintinline{bash}{I__employee} interface.}
	\label{fig:run-time-information-gathering-get-field}
\end{figure}