### 2019-08-11 chapters/55-related.tex

\chapter{Related Work}\label{chap:relatedwork}
Give a brief overview of the work relevant for your thesis.
Closely related work should be compared in depth: what are the pros and cons of their work compared to yours?

\todo{dts-gen package: https://github.com/microsoft/dts-gen}