### 2019-08-11 figures/approach/typescript-declaration-files-generation-method/examples/templates/module-class-template.tex

\begin{figure}[h]
	\centering
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.48\textwidth}
			\jscode{code/typescript-declaration-file-generation/templates/module-class/index.js}
		\end{minipage}
	\end{lrbox}
	\subfloat[index.js - Entry point to be executed with \mintinline{bash}{node index.js}]{\usebox{\mintedbox}}
	\hfill
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.5\textwidth}
			\jscode{code/typescript-declaration-file-generation/templates/module-class/calculator.js}
		\end{minipage}
	\end{lrbox}
	\subfloat[calculator.js - Module for which the TS Declaration File is going to be generated]{\usebox{\mintedbox}}

	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.7\textwidth}
			\tscode{code/typescript-declaration-file-generation/templates/module-class/calculator.d.ts}
		\end{minipage}
	\end{lrbox}
	\subfloat[calculator/index.d.ts]{\usebox{\mintedbox}}

	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.7\textwidth}
			\tscode{code/typescript-declaration-file-generation/templates/module-class/index.ts}
		\end{minipage}
	\end{lrbox}
	\subfloat[index.ts - TypeScript file using the generated declaration file]{\usebox{\mintedbox}}
	\caption[module-class.d.ts template example]{\textbf{module-class.d.ts template example} - Declaration file for the \mintinline{bash}{calculator} module is generated using the \mintinline{bash}{module-class.d.ts} template. Exported module can be used as a constructor within a TypeScript project.}
	\label{fig:ts-declaration-file-module-class-template}
\end{figure}