### 2019-08-11 figures/approach/typescript-declaration-files-generation-method/examples/templates/module-function-template.tex

\begin{figure}[h]
	\centering
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.9\textwidth}
			\jscode{code/typescript-declaration-file-generation/templates/module-function/index.js}
		\end{minipage}
	\end{lrbox}
	\subfloat[index.js - Entry point to be executed with \mintinline{bash}{node index.js}]{\usebox{\mintedbox}}

	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.9\textwidth}
			\jscode{code/typescript-declaration-file-generation/templates/module-function/snake-to-camel.js}
		\end{minipage}
	\end{lrbox}
	\subfloat[snake-to-camel.js - Module for which the TS Declaration File is going to be generated]{\usebox{\mintedbox}}

	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.8\textwidth}
			\tscode{code/typescript-declaration-file-generation/templates/module-function/index.d.ts}
		\end{minipage}
	\end{lrbox}
	\subfloat[snake-to-camel/index.d.ts]{\usebox{\mintedbox}}

	\caption[module-function.d.ts template example]{\textbf{module-function.d.ts template example} - Declaration file for the \mintinline{bash}{snake-to-camel} module is generated using the \mintinline{bash}{module-function.d.ts} template.}
	\label{fig:ts-declaration-file-module-function-template}
\end{figure}