### 2019-08-13 figures/approach/run-time-information-gathering/analysis/input-value.tex

\begin{figure}[tp]
	\centering
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.4\textwidth}
			\jscode{code/run-time-information-gathering/simple-examples/input-value.js}
		\end{minipage}
	\end{lrbox}
	\subfloat[Instrumented JS code]{\usebox{\mintedbox}}
	\hfill
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.58\textwidth}
			\jsoncode{code/run-time-information-gathering/simple-examples/output-input-value.json}
		\end{minipage}
	\end{lrbox}
	\subfloat[Run-time information output]{\usebox{\mintedbox}}
	\caption[\textit{inputValue} interaction example]{\textbf{\textit{inputValue} interaction example} - The type of the arguments of function \textit{foo()} is recorded. Several additional fields and entries of the final output JSON file were omitted for simplicity.}
	\label{fig:run-time-information-gathering-input-value}
\end{figure}