### 2019-08-16 figures/experiments/results/datadog-metrics.tex

\begin{figure}[tp]
	\centering
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.49\textwidth}
			\tscode{code/experiments/results/module/datadog-metrics/generated.d.ts}
		\end{minipage}
	\end{lrbox}
	\subfloat[datadog-metrics/index.d.ts - Generated]{\usebox{\mintedbox}}
	\hfill
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.49\textwidth}
			\tscode{code/experiments/results/module/datadog-metrics/definitely-typed.ts}
		\end{minipage}
	\end{lrbox}
	\subfloat[datadog-metrics/index.d.ts - DefinitelyTyped]{\usebox{\mintedbox}}

	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.95\textwidth}
			\tscode{code/experiments/results/module/datadog-metrics/logger-datadog-metrics.js}
		\end{minipage}
	\end{lrbox}
	\subfloat[datadog-metrics/logger.js]{\usebox{\mintedbox} \label{fig:subfloat-datadog-logger} }

	\caption[Missing properties | datadog-metrics]{\textbf{Missing properties | datadog-metrics} - Properties \mintinline{text}{aggregator} and \mintinline{text}{reporter} are not in the DefinitelyTyped version, but they appear in the generated declaration file. However, they are indeed used by the library, as exposed in lines 2 and 3 of the library's source code shown in \protect\subref{fig:subfloat-datadog-logger}.}
	\label{fig:experiments-results-module-datadog-metrics}
\end{figure}