### 2019-08-16 figures/experiments/results/module-class-results-route-parser.tex

\begin{figure}[tp]
	\centering
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.8\textwidth}
			\tscode{code/experiments/results/module-class/route-parser/generated.d.ts}
		\end{minipage}
	\end{lrbox}
	\subfloat[route-parser/index.d.ts - Generated]{\usebox{\mintedbox}}

	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.8\textwidth}
			\tscode{code/experiments/results/module-class/route-parser/definitely-typed.d.ts}
		\end{minipage}
	\end{lrbox}
	\subfloat[route-parser/index.d.ts - DefinitelyTyped]{\usebox{\mintedbox}}

	\caption[Module-class results | route-parser]{\textbf{Module-class results | route-parser} - Constructor and methods are correctly identified. Specific signature of dictionary \mintinline{text}{params} is not detected. Instead, a generic \mintinline{text}{object} is inferred.}
	\label{fig:experiments-results-module-class-route-parser}
\end{figure}