### 2019-08-16 figures/experiments/results/module-class-results-timer-machine.tex

\begin{figure}[tp]
	\centering
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.8\textwidth}
			\tscode{code/experiments/results/module-class/timer-machine/generated.d.ts}
		\end{minipage}
	\end{lrbox}
	\subfloat[timer-machine/index.d.ts - Generated]{\usebox{\mintedbox}}

	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.8\textwidth}
			\tscode{code/experiments/results/module-class/timer-machine/definitely-typed.d.ts}
		\end{minipage}
	\end{lrbox}
	\subfloat[timer-machine/index.d.ts - DefinitelyTyped]{\usebox{\mintedbox}}

	\caption[Module-class results | timer-machine]{\textbf{Module-class results | timer-machine} - Parameter \mintinline{text}{started} is inferred as \mintinline{text}{undefined} instead of marking it as optional. Methods that were not executed do not appear in the generated declaration file.}
	\label{fig:experiments-results-module-class-timer-machine}
\end{figure}