### 2019-08-16 figures/experiments/results/module-results.tex

\begin{figure}[tp]
	\centering
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.8\textwidth}
			\tscode{code/experiments/results/module/is-uuid/generated.d.ts}
		\end{minipage}
	\end{lrbox}
	\subfloat[is-uuid/index.d.ts - Generated]{\usebox{\mintedbox}}

	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.8\textwidth}
			\tscode{code/experiments/results/module/is-uuid/definitely-typed.d.ts}
		\end{minipage}
	\end{lrbox}
	\subfloat[is-uuid/index.d.ts - DefinitelyTyped]{\usebox{\mintedbox}}

	\caption[Module results | is-uuid]{\textbf{Module results | is-uuid} - Methods that were not executed are not included in the declaration file.}
	\label{fig:experiments-results-module-is-uuid}
\end{figure}