### 2019-08-16 figures/experiments/template-distribution-definitely-typed.tex

\begin{figure}[tp]
	\centering
	\begin{tikzpicture}
		\begin{axis}[
			ybar,
			width=0.7\textwidth,
			ybar=0pt,
			ymax=6000,
			enlargelimits=0.15,
			bar width=0.08\textwidth,
			legend style={at={(0.5,-0.2)}, anchor=north,legend columns=-1},
			symbolic x coords={module-class,module-function,module},
			xtick=data,
			nodes near coords, 
			nodes near coords align={vertical},
			x tick label style={rotate=45,anchor=east},
		]
		\addplot coordinates {
			(module-class, 503)
			(module-function, 1201)
			(module, 4379)
		};
		\end{axis}
	\end{tikzpicture}

	\caption[TypeScript templates distribution in DefinitelyTyped]{\textbf{TypeScript templates distribution in DefinitelyTyped} - Out of a total of 6029, 72\% of the modules uploaded to the DefinitelyTyped repository use the \mintinline{text}{module} template.
	}

	\label{fig:experiments-typescript-templates-distribution-definitely-typed}
\end{figure}