### 2019-08-20 chapters/3-background.tex
Complete Jalangi section

\chapter{Background}\label{chap:background}
\section{JavaScript} \label{sec:background-javascript}
JavaScript (abbreviated JS) is the standard programming language for developing client-side web applications. Together with HTML and CSS, they conform the key technology stack of the Web.

JavaScript is compliant with the ECMAScript Language Specification \citep{ecma-script}. It is an interpreted, dynamically typed language that allows to program small scripts very fast.  Its low learning curve, the possibility of fast prototyping and its easy interaction with the browser's DOM made it the ideal programming languages for introducing dynamic behavior to web applications.

Numerous JavaScript frameworks that aid the development of web applications have been created over the years. \mintinline{text}{jQuery} is used by 74.2\% of the 10 million most popular websites, making it the most popular JavaScript framework in the market \citep{jquery}\citep{w3techs-javascript-libraries-statistics}. Other frameworks like \mintinline{text}{AngularJS}, \mintinline{text}{React} and \mintinline{text}{Vue.js} appeared as an elegant alternative for building complex and scalable web applications \citep{angularjs}\citep{react}\citep{vuejs}. A very simple example of how to use \mintinline{text}{jQuery} for adding dynamic behavior to a web page is shown in \coderef{code:background-jquery-example}.

\begin{code}
	\htmlcode{code/background/javascript/jquery-example.html}
	\captionsetup{aboveskip=0pt, belowskip=10pt}
	\caption[Example using JavaScript and HTML]{\textbf{Example using JavaScript and HTML} - Very simple example that uses library jQuery to hide an element from the DOM when it gets clicked. The example was extracted from W3Schools website.}
	\label{code:background-jquery-example}
\end{code}

All browsers have built-in interpreters that support running JavaScript code. The language is also used to develop backend applications that run in server side runtime environments like NodeJS \citep{nodejs}.

\figref{fig:background-survey-programming-languages} shows that it is the most popular programming languages among professional developers. Additionally, 5 of the 12 most popular frameworks used for web developing are JavaScript frameworks. Similarly, it can be seen in \figref{fig:background-programming-languages-evolution} that JavaScript's popularity remained constant in the last 3 years while being the most popular programming language for seven consecutive years.

\input{figures/background/stackoverflowSurvey-ProgrammingLanguages2019.tex}

\input{figures/background/stackoverflowSurvey-ProgrammingLanguagesEvolution.tex}

\subsection{Types}
Section 8 of the ECMAScript Language Specification presents the allowed types for a variable \citep{ecma-script}. These are: \mintinline{text}{undefined}, \mintinline{text}{null}, \mintinline{text}{boolean}, \mintinline{text}{string}, \mintinline{text}{number}, \mintinline{text}{object} and \mintinline{text}{symbol} for ES6/ECMAScript 2015. All of them except for \mintinline{text}{object} are called primitives.

It must be noted that functions and arrays are not built-in types. Functions are considered to be `member of the Object type that is an instance of the standard built-in Function constructor and that may be invoked as a subroutine' \citep{ecma-script}. The \mintinline{text}{Function} object is a built-in object with a \mintinline{text}{[[Call]]} internal property, which means that it can be invoked. Similarly, arrays are instances of the built-in \mintinline{text}{Array} object. Both concepts are shown in \coderef{code:background-functions-and-arrays}.

\begin{code}
	\jscode{code/background/javascript/functions-and-arrays.js}
	\captionsetup{aboveskip=0pt, belowskip=10pt}
	\caption[Functions and arrays are built-in objects in JS]{\textbf{Functions and arrays are built-in objects in JS} - A function can be created using the \mintinline{text}{function} keyword or using the \mintinline{text}{Function} built-in constructor. Similarly, arrays can be created using brackets or the \mintinline{text}{Array} built-in object.}
	\label{code:background-functions-and-arrays}
\end{code}

Finally, the \mintinline{text}{typeof} operator is not an exact match with the built-in types. It surprisingly returns \mintinline{text}{'object'} for \mintinline{text}{null}, \mintinline{text}{'function'} for a \mintinline{text}{function} and \mintinline{text}{'object'} for an \mintinline{text}{array}. \coderef{code:background-javascript-typeof} gives some examples about this operator.

\begin{code}
	\jscode{code/background/javascript/javascript-typeof.js}
	\captionsetup{aboveskip=0pt, belowskip=10pt}
	\caption[typeof JavaScript operator]{\textbf{\mintinline{text}{typeof} JavaScript operator} - Examples of the \mintinline{text}{typeof} operator for \mintinline{text}{null}, \mintinline{text}{function} and \mintinline{text}{array}.}
	\label{code:background-javascript-typeof}
\end{code}

\subsection{Wrapper Objects for Primitives}
JavaScript also includes the built-in objects \mintinline{text}{Boolean}, \mintinline{text}{String} and \mintinline{text}{Number}. Whenever a method or a property of a primitive value of type \mintinline{text}{boolean}, \mintinline{text}{string} and \mintinline{text}{number} is accessed, JavaScript will convert the primitive value into its corresponding built-in object before accessing the property or invoking the method, as illustrated in \coderef{code:background-javascript-wrapper-objects}. It is however considered a bad practice to explicitly use the built-in objects instead of the primitive values. Several JavaScript linters like JSHint will consider this an error \citep{jshint-wrapper-objects-error}.

\begin{code}
	\jscode{code/background/javascript/javascript-wrapper-objects.js}
	\captionsetup{aboveskip=0pt, belowskip=10pt}
	\caption[Wrapper Objects for primitives]{\textbf{Wrapper Objects for primitives} - When a method is invoked on a primitive value like \mintinline{text}{number} or \mintinline{text}{string}, an intermediate wrapper object is created.}
	\label{code:background-javascript-wrapper-objects}
\end{code}

\subsection{TypeCoercion} \label{sec:background-js-type-coercion}
Type coercion is the process of converting one type into another one. In JavaScript, a variable can be converted into different types depending on the operator and the value of the other operands. The specific definition can be found in the ECMAScript Language Specification \citep{ecma-script}. The following section will cover the basic forms of type coercion but providing a detailed explanation of how JavaScript converts a variable type into another one for each operator is beyond the scope of this section. It is however intended to give insight about this kind of behavior to show its importance while performing type inference in JavaScript. As a result, specific operator related type coercion is explained only for operator \mintinline{text}{+}.

Finally, there are excellent references regarding JavaScript's Type Coercion, like Kyle Simpson's book `You Don't Know JS: Types \& Grammar' \citep{you-dont-know-js}.

\subsubsection{Types of conversion}
\label{types_of_conversion}
There are two types of type coercion: explicit and implicit \citep{you-dont-know-js}. Explicit type coercion is simply when the developer chooses to convert one type into another one. The developer achieves this by writing a piece of code that shows her intention, as shown in \coderef{code:background-explicit-type-coercion}. This type of coercion is also called type casting \citep{you-dont-know-js}.

\begin{code}
	\jscode{code/background/javascript/type-coercion/explicit-type-coercion.js}
	\captionsetup{aboveskip=0pt, belowskip=10pt}
	\caption[Explicit JavaScript Type Coercion]{\textbf{Explicit JavaScript Type Coercion} - The developer explicitly transforms a type into another one. The return values of \mintinline{text}{String()}, \mintinline{text}{Number()} and \mintinline{text}{Boolean()} are always of type \mintinline{text}{string}, \mintinline{text}{number} and \mintinline{text}{boolean}, respectively.}
	\label{code:background-explicit-type-coercion}
  \end{code}

On the other hand, depending on the operator and the context, JavaScript will perform implicit type transformations automatically. If the developer is not aware of this, she may face some unexpected behavior, as shown in \coderef{code:background-implicit-type-coercion}.

\begin{code}
	\jscode{code/background/javascript/type-coercion/implicit-type-coercion.js}
	\captionsetup{aboveskip=0pt, belowskip=10pt}
	\caption[Implicit JavaScript Type Coercion]{\textbf{Implicit JavaScript Type Coercion} - Examples given by Douglas Crockford in his talk `JavaScript: The Good Parts' at Google\footnote{https://www.youtube.com/watch?v=hQVTIJBZook}.}
	\label{code:background-implicit-type-coercion}
\end{code}

In any case, JavaScript will always convert any value into a primitive value \citep{ecma-script}, regardless of explicit or implicit type coercion. There is no coercion mechanism that results in an \mintinline{text}{object} or a \mintinline{text}{function}.

It is fundamental to understand how a value gets coerced into one of these types. JavaScript mainly performs this by applying one of three so called \mintinline{text}{Abstract Operations} \citep{ecma-script}: \mintinline{text}{ToString()}, \mintinline{text}{ToNumber()} and \mintinline{text}{ToBoolean()}.

The operation \mintinline{text}{ToPrimitive()} is also called within the operators. However, each operator will call either \mintinline{text}{ToString()}, \mintinline{text}{ToNumber()} or \mintinline{text}{ToBoolean()} after calling \mintinline{text}{ToPrimitive()}. This operator will be explained in an independent section.

\subsubsection{ToString}
JavaScript performs this conversion in a very intuitive way, as explained in \coderef{code:background-to-string-operation}. Every primitive gets converted as expected. Objects are converted into strings by calling \mintinline{text}{ToPrimtive()} with \mintinline{text}{hint = string}, as presented in \coderef{code:background-to-string-implementation}.

\begin{code}
	\jscode{code/background/javascript/type-coercion/to-string.js}
	\captionsetup{aboveskip=0pt, belowskip=10pt}
	\caption[ToString implementation]{\textbf{ToString implementation}}
	\label{code:background-to-string-implementation}
\end{code}

\begin{code}
	\jscode{code/background/javascript/type-coercion/string-conversion.js}
	\captionsetup{aboveskip=0pt, belowskip=10pt}
	\caption[ToString operation]{\textbf{ToString operation}}
	\label{code:background-to-string-operation}
\end{code}

\subsubsection{ToNumber}
\coderef{code:background-to-number-operation} presents this type of conversion. Strings and boolean values get converted in the expected way. Moreover, \mintinline{text}{undefined} will be converted into \mintinline{text}{NaN} and \mintinline{text}{null} into \mintinline{text}{0}. Similarly, objects are converted into numbers by calling \mintinline{text}{ToPrimtive()} with \mintinline{text}{hint = number}, as presented in \coderef{code:background-to-number-implementation}.

\begin{code}
	\jscode{code/background/javascript/type-coercion/to-number.js}
	\captionsetup{aboveskip=0pt, belowskip=10pt}
	\caption[ToNumber implementation]{\textbf{ToNumber implementation}}
	\label{code:background-to-number-implementation}
\end{code}

\begin{code}
	\jscode{code/background/javascript/type-coercion/number-conversion.js}
	\captionsetup{aboveskip=0pt, belowskip=10pt}
	\caption[ToNumber operation]{\textbf{ToNumber operation}}
	\label{code:background-to-number-operation}
\end{code}

\subsubsection{ToBoolean}
Only the values presented in \coderef{code:background-to-boolean-operation} are converted into \mintinline{text}{false}. Everything else is coerced to \mintinline{text}{true}. That means, that every non-empty \mintinline{text}{string}, every non-zero \mintinline{text}{number} and every \mintinline{text}{object} gets coerced to \mintinline{text}{true}.

Values that are converted into \mintinline{text}{false}, like \mintinline{text}{null} or \mintinline{text}{undefined} are called \textit{falsy} values \citep{you-dont-know-js}.

\begin{code}
	\jscode{code/background/javascript/type-coercion/boolean-conversion.js}
	\captionsetup{aboveskip=0pt, belowskip=10pt}
	\caption[ToBoolean operation]{\textbf{ToBoolean operation}}
	\label{code:background-to-boolean-operation}
\end{code}

\subsubsection{ToPrimitive}
The signature of this function is \mintinline{text}{ToPrimitive(value, preferredType)}. The allowed values for \mintinline{text}{preferredType} are \mintinline{text}{string} or \mintinline{text}{number}. Every JavaScript operator will try to convert an object into a primitive by invoking \mintinline{text}{ToPrimitive} with a different value for \mintinline{text}{preferredType}. If no value is passed, \mintinline{text}{number} is considered as the default, except for JavaScript native \mintinline{text}{Date} objects, where \mintinline{text}{string} is used as default.

An implementation of this behavior is shown in \coderef{code:background-to-primitive-operation}.

\begin{code}
	\jscode{code/background/javascript/type-coercion/to-primitive/to-primitive.js}
	\captionsetup{aboveskip=0pt, belowskip=10pt}
	\caption[ToPrimitive operation]{\textbf{ToPrimitive operation}}
	\label{code:background-to-primitive-operation}
\end{code}

An important aspect that should be considered, is that the type of the return value of \mintinline{text}{ToPrimitive()} does not necessarily have to match the chosen \mintinline{text}{preferredType}. This means concretely that, for example, you can call \mintinline{text}{ToPrimitive(a, number)} and get a \mintinline{text}{string} as a result. Of course, when performing an explicit type conversion, JS will eventually convert whatever \mintinline{text}{ToPrimitive} returns into the correct type. But some operators will perform different implicit type conversions \textit{after} calling \mintinline{text}{ToPrimitive()}. The specific type conversion will then be dependent on the type of the return value of \mintinline{text}{ToPrimitive()} and not on the chosen \mintinline{text}{preferredType}.

All objects that are converted to \mintinline{text}{boolean} will be coerced to \mintinline{text}{true}. The following paragraphs explain how objects get converted into strings or numbers.

\subsubsection{Object to String}
The object is first converted into a primitive value calling the \mintinline{text}{ToPrimitive()} method, using \mintinline{text}{string} as \mintinline{text}{preferredType}. The obtained primitive value is then normally converted into a \mintinline{text}{string}, as explained before in \coderef{code:background-to-string-implementation}.

\coderef{code:background-object-into-string} shows a straightforward example. On the other hand, \coderef{code:background-object-into-string-not-string-return-value} provides an example where \mintinline{text}{ToPrimitive(val, String)} does not return a \mintinline{text}{string}, either because simply \mintinline{text}{toString()} does not return a \mintinline{text}{string} or because \mintinline{text}{valueOf()} gets being called instead of \mintinline{text}{toString()}.

\begin{code}
	\jscode{code/background/javascript/type-coercion/to-primitive/normal-object-to-string.js}
	\captionsetup{aboveskip=0pt, belowskip=10pt}
	\caption[Object into string conversion]{\textbf{Object into string conversion} - An object has a \mintinline{text}{toString()} method that returns a string.}
	\label{code:background-object-into-string}
\end{code}

\begin{code}
	\jscode{code/background/javascript/type-coercion/to-primitive/object-to-string-returning-not-a-string.js}
	\captionsetup{aboveskip=0pt, belowskip=10pt}
	\caption[Object into string conversion]{\textbf{Object into string conversion} - An object that does not return a \mintinline{text}{string} even though \mintinline{text}{ToPrimitive()} is called with \mintinline{text}{hint = string}.}
	\label{code:background-object-into-string-not-string-return-value}
\end{code}

\subsubsection{Object to Number}
The object gets converted into a primitive by calling \mintinline{text}{ToPrimitive(val, Number)}. Normal number conversion is then applied to this primitive value. \coderef{code:background-object-into-number} provides a basic example. Similarly, \coderef{code:background-object-into-string-not-number-return-value} shows how \mintinline{text}{ToPrimitive()} might return something that is not a \mintinline{text}{number}.

\begin{code}
	\jscode{code/background/javascript/type-coercion/to-primitive/normal-object-to-number.js}
	\captionsetup{aboveskip=0pt, belowskip=10pt}
	\caption[Object into number conversion]{\textbf{Object into number conversion} - An object has a \mintinline{text}{valueOf()} method that returns a \mintinline{text}{number}.}
	\label{code:background-object-into-number}
\end{code}

\begin{code}
	\jscode{code/background/javascript/type-coercion/to-primitive/object-to-number-returning-not-a-number.js}
	\captionsetup{aboveskip=0pt, belowskip=10pt}
	\caption[Object into number conversion]{\textbf{Object into number conversion} - An object that does not return a \mintinline{text}{number} even though \mintinline{text}{ToPrimitive()} is called with \mintinline{text}{hint = number}.}
	\label{code:background-object-into-string-not-number-return-value}
\end{code}

\subsubsection{The \mintinline{text}{+} operator}
If any of the operands is an object, it will first convert it into a primitive by passing no \mintinline{text}{preferredType} to the \mintinline{text}{ToPrimitive()} method. Afterwards, if any value is a \mintinline{text}{string}, this operator will implicitly convert both values into a \mintinline{text}{string} and perform a normal string concatenation between them. Else, it will implicitly convert each value into a \mintinline{text}{number} and perform a normal arithmetic addition. An implementation is presented in \coderef{code:background-plus-operator-implementation}.

\coderef{code:background-plus-operator-simple-examples} shows some basic examples. As explained before and shown in \coderef{code:background-plus-operator-object-example}, \mintinline{text}{ToPrimitive()} may return a \mintinline{text}{string}, even when it is called with no \mintinline{text}{preferredType}. This can happen because:
\begin{enumerate}
	\item \mintinline{text}{valueOf()} is returning a String.
	\item \mintinline{text}{valueOf()} is not returning a primitive, which would make \mintinline{text}{toString()} being called instead.
	\item One of the operands is the native JavaScript \mintinline{text}{Date} Object.
\end{enumerate}

\begin{code}
	\jscode{code/background/javascript/type-coercion/plus-operator-implementation.js}
	\captionsetup{aboveskip=0pt, belowskip=10pt}
	\caption[JavaScript + operator implementation]{\textbf{JavaScript \mintinline{text}{+} operator implementation}}
	\label{code:background-plus-operator-implementation}
\end{code}

\begin{code}
	\jscode{code/background/javascript/type-coercion/plus-operator-simple-examples.js}
	\captionsetup{aboveskip=0pt, belowskip=10pt}
	\caption[JavaScript + operator examples]{\textbf{JavaScript \mintinline{text}{+} operator examples}}
	\label{code:background-plus-operator-simple-examples}
\end{code}

\begin{code}
	\jscode{code/background/javascript/type-coercion/plus-operator-object.js}
	\captionsetup{aboveskip=0pt, belowskip=10pt}
	\caption[JavaScript + operator with object]{\textbf{JavaScript \mintinline{text}{+} operator with object}}
	\label{code:background-plus-operator-object-example}
\end{code}

\subsection{NPM Packages} \label{sec:background-npm-packages}
The Node Package Manager (NPM) is the default package manager for NodeJS, a JavaScript runtime environment built on Google Chrome's V8 JavaScript engine \citep{nodejs}. It offers a public registry where JavaScript libraries can be uploaded to be used by the developer's community.

Dependencies can be installed either globally or as local dependencies of specific a JavaScript projects. NPM installs the dependencies described in the \mintinline{text}{package.json} file and local dependencies are saved in the \mintinline{text}{node_modules} directory in the project's root directory. An example of NPM's usage is shown in \coderef{code:background-npm-usage-example}. The registry contains currently more than 1 million modules and it has 11000 million downloads per week.

\begin{code}
	\begin{bashinline}
$ cat index.js 
const abs = require('abs');
console.log(abs('/foo'));

$ node index.js
Error: Cannot find module 'abs'

$ npm i abs
$ node index.js
/foo
	  \end{bashinline}
	\caption[NPM usage example]{\textbf{NPM usage example} - Requiring the module before installing the dependencies will fail since the library is not available. After installing the dependencies locally under the \mintinline{text}{node_modules} directory, the \mintinline{text}{abs} module can be imported using the \mintinline{text}{require} function.}
	\label{code:background-npm-usage-example}
  \end{code}

\section{TypeScript} \label{sec:background-typescript}
\subsection{Declaration Files} \label{sec:declaration-files-background}
Declaration files contain a typed description of an external JavaScript library's API. It allows the IDE to use code intelligence features like auto-completion and avoids misusing the library by declaring specific interfaces and types required by the library. They are necessary for integrating new TypeScript projects with existing JavaScript libraries that are still developed in plain JavaScript.

Declaration files have the extension \mintinline{text}{.d.ts}. \figref{fig:background-declaration-files-calculator-example} provides an example where a new TypeScript program is using the declaration file of an external JavaScript library called \mintinline{text}{calculator}. The example  library \mintinline{text}{calculator} is not part of the TypeScript project, it is imported in run-time after compilation when the JavaScript code is executed.

\input{figures/background/typescript/declaration-files/calculator-example.tex}

It is important to mention that the compiler and the IDE will blindly trust the declaration file. There are no run-time checks being performed on the declaration file and the JavaScript code itself. TypeScript is not aware if the JavaScript library corresponding to the declaration file even exists. \figref{fig:background-declaration-files-calculator-error-example} exposes a scenario where a declaration file that does not match the JavaScript implementation. The declaration file states that types \mintinline{text}{number} and \mintinline{text}{string} are allowed for \mintinline{text}{sum} method but the JavaScript code explicitly throws an error when the arguments are not of type \mintinline{text}{number}. The IDE, however, will provide suggestions based on the inaccurate declaration file. The compiler itself will use the declaration file to perform type checking. Therefore, the code is compiled without errors and the error is only encountered at run-time.

\input{figures/background/typescript/declaration-files/calculator-example-error.tex}

It is clear that inaccurate declaration files will cause errors that are tedious and difficult to track and debug. Discrepancies with the actual JavaScript implementation is going to frustrate developers, stopping them from using TypeScript as an alternative to JavaScript. Hence the importance of accurate declaration files that match the JavaScript implementation.

There are two possibilities for importing declaration files of NPM packages:
\begin{itemize}
	\item If the JavaScript library is written in TypeScript, then the declaration file will be generated automatically by the compiler using the \mintinline{text}{--declaration} flag. The generated declaration file is bundled with the JavaScript distribution files and uploaded to the NPM registry as a part of the package. The \mintinline{text}{types} field in the \mintinline{text}{package.json} file should point to the declaration file. Consequently, the declaration file and the JavaScript code will always be synchronized.
	\item If the library is not written in TypeScript, the declaration files have to be manually created and uploaded to the DefinitelyTyped repository \citep{definitely-typed-repository}. After a successful pull request, files will be automatically published to the \mintinline{text}{@types} organization on NPM \citep{types-organization-npm}.
\end{itemize}

\todo{Agregar templates!}

\subsection{Definitely Typed}
The declaration files for NPM packages that do not already include them in the package are uploaded to the DefinitelyTyped repository \citep{definitely-typed-repository}. It currently contains declaration files for more than 6000 modules. The developers community adds new declaration files or updates existing ones by creating a pull request with the changes. If it is approved, the changes will be automatically pushed to the NPM registry through Microsoft's \mintinline{text}{types-publisher} tool \citep{types-publisher}.

\subsubsection{Consumption}
As shown in \coderef{code:background-definitely-typed-consumption}, the declaration files can be simply imported by running \mintinline{text}{npm install --save-dev @types/THE-MODULE}. NPM will install the files under the \mintinline{text}{node_modules/@types} directory and the compiler will include the declaration files automatically.

\begin{code}
	\begin{bashinline}
$ npm install --save-dev @types/abs 
$ cat node_modules/@types/abs/index.d.ts

// Type definitions for abs 1.3
// Project: https://github.com/ionicabizau/abs
// Definitions by: Aya Morisawa <https://github.com/AyaMorisawa>
// Definitions: https://github.com/DefinitelyTyped/DefinitelyTyped

/**
 * Compute the absolute path of an input.
 * @param input The input path.
 */
declare function Abs(input: string): string;
export = Abs;
	  \end{bashinline}
	\caption[DefinitelyTyped files consumption]{\textbf{DefinitelyTyped files consumption} - The declaration files are installed as a common NPM package under the \mintinline{text}{@types} directory.}
	\label{code:background-definitely-typed-consumption}
  \end{code}

\section{Jalangi} \label{sec:jalangi}
Jalangi is a framework for performing dynamic analysis in JavaScript with configurable modules \citep{DBLP:conf/sigsoft/SenKBG13}. It first instruments the code and then executes the instrumented code adding the analysis module. Jalangi offers different callbacks that match virtually every JavaScript operations. The relevant callbacks for this work are listed below with short descriptions that have been directly copied from Jalangi's documentation \citep{jalangi-docs}:
\begin{itemize}
	\item \textbf{binaryPre()}: `This callback is called before a binary operation.'
	\item \textbf{declare()}: `This callback is triggered at the beginning of a scope for every local variable declared in the scope, for every formal parameter, for every function defined using a function statement, for arguments variable, and for the formal parameter passed in a catch statement.'
	\item \textbf{getField()}: `This callback is called after a property of an object is accessed.'
	\item \textbf{putFieldPre()}: `This callback is called before a property of an object is written.'
	\item \textbf{functionEnter()}: `This callback is called before the execution of a function body starts.'
	\item \textbf{functionExit()}: `This callback is called when the execution of a function body completes'
	\item \textbf{invokeFun()}: `This callback is called after a function, method, or constructor invocation.'
	\item \textbf{invokeFunPre()}: `This callback is called before a function, method, or constructor invocation.'
	\item \textbf{unaryPre()}: `This callback is called before a unary operation.'
	\item \textbf{write()}: `This callback is called before a variable is written.'
  \end{itemize}

\figref{fig:background-jalangi} provides an example for the usage of Jalangi where invocations of a function are detected and printed by stdout.

\input{figures/background/typescript/jalangi/jalangi-example.tex}