### 2019-08-20 figures/background/typescript/jalangi/jalangi-example.tex

\begin{figure}[tp]
	\centering
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.34\textwidth}
			\jscode{code/background/jalangi/sum.js}
		\end{minipage}
	\end{lrbox}
	\subfloat[sum.js]{\usebox{\mintedbox}}
	\hfill
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.64\textwidth}
			\jscode{code/background/jalangi/analysis.js}
		\end{minipage}
	\end{lrbox}
	\subfloat[analysis.js]{\usebox{\mintedbox}}

	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.8\textwidth}
			\begin{bashinline}
$ node jalangi.js --analysis analysis.js sum.js
Function sum was invoked with params (1,2) and returned 3
Function sum was invoked with params (3,4) and returned 7
Function sum was invoked with params (5,6) and returned 11
			\end{bashinline}
		\end{minipage}
	\end{lrbox}
	\subfloat[Console output]{\usebox{\mintedbox}}

	\caption[Jalangi usage example]{\textbf{Jalangi usage example} - \mintinline{text}{sum.js} file is instrumented and executed alongside the analysis module defined in \mintinline{text}{analysis.js}.}
	\label{fig:background-jalangi}
\end{figure}