### 2019-08-22 figures/background/typescript/classes/class-example.tex

\begin{figure}[tp]
	\centering
	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.9\textwidth}
			\tscode{code/background/typescript/classes/index.ts}
		\end{minipage}
	\end{lrbox}
	\subfloat[index.ts]{\usebox{\mintedbox}}

	\begin{lrbox}{\mintedbox}
		\begin{minipage}{0.9\textwidth}
			\jscode{code/background/typescript/classes/index.js}
		\end{minipage}
	\end{lrbox}
	\subfloat[index.js]{\usebox{\mintedbox}}

	\caption[TypeScript classes compilation]{\textbf{TypeScript classes compilation} - A TypeScript class is compiled into the common ES5 construct used for emulating classes in JavaScript. Private and public methods are only checked at compile-time. Both methods are exported in the same way, since JavaScript has no concept of method's visibility.}
	\label{fig:background-typescript-class-example}
\end{figure}