### 2020-01-04 ecoop-2020/main.tex
cosmetics and further tweaking

\documentclass[a4paper,english,cleveref, autoref]{lipics-v2019}
% ECOOP 2020: 25 pages excluding references
%This is a template for producing LIPIcs articles. 
%See lipics-manual.pdf for further information.
%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
%for section-numbered lemmas etc., use "numberwithinsect"
%for enabling cleveref support, use "cleveref"
%for enabling cleveref support, use "autoref"

\usepackage{pgfplots}
\usepackage{tikz}
\usetikzlibrary{positioning}

\input{macros}

\definecolor{backgroundgray}{rgb}{.94,.94,.94}
\definecolor{lightgray}{rgb}{.35,.35,.5}
\definecolor{darkgray}{rgb}{.2, .2, .2}
\definecolor{purple}{rgb}{0.65, 0.12, 0.82}

\newcommand{\secref}[1]{Section~\ref{#1} - \nameref{#1}}
\newcommand{\figref}[1]{Figure~\ref{#1}}
\newcommand{\coderef}[1]{Listing~\ref{#1}}

\lstdefinelanguage{JavaScript}{
  keywords={typeof, new, true, false, catch, function, return, null, catch, switch, var, if, in, while, do, else, case, break, undefined},
  keywordstyle=\color{black}\bfseries,
  ndkeywords={class, export, boolean, throw, implements, import, this},
  ndkeywordstyle=\color{black}\bfseries,
  identifierstyle=\color{darkgray},
  sensitive=false,
  comment=[l]{//},
  morecomment=[s]{/*}{*/},
  commentstyle=\color{lightgray}\slshape,
  stringstyle=\color{darkgray}\ttfamily,
  morestring=[b]',
  morestring=[b]"
}

\lstdefinelanguage{TypeScript}{
  keywords={typeof, new, true, false, catch, function, return, null, catch, switch, var, if, in, while, do, else, case, break, undefined},
  keywordstyle=\color{black}\bfseries,
  ndkeywords={class, export, boolean, throw, implements, import, this, declare, constructor, namespace, interface},
  ndkeywordstyle=\color{black}\bfseries,
  identifierstyle=\color{darkgray},
  sensitive=false,
  comment=[l]{//},
  morecomment=[s]{/*}{*/},
  commentstyle=\color{lightgray}\slshape,
  stringstyle=\color{darkgray}\ttfamily,
  morestring=[b]',
  morestring=[b]"
}

\lstset{
   backgroundcolor=\color{backgroundgray},
   extendedchars=true,
   basicstyle=\footnotesize\ttfamily,
   showstringspaces=false,
   showspaces=false,
   numbers=left,
   numberstyle=\footnotesize,
   numbersep=9pt,
   tabsize=2,
   breaklines=true,
   showtabs=false
}

\graphicspath{{./graphics/}}%helpful if your graphic files are in another directory

\bibliographystyle{plainurl}% the mandatory bibstyle

\title{Generation of TypeScript Declaration Files from JavaScript Code}

% \titlerunning{Dummy short title}%optional, please use if title is longer than one line

\author{John Q. Public}{Dummy University Computing Laboratory, Country \and My second affiliation, Country \and \url{http://www.myhomepage.edu} }{johnqpublic@dummyuni.org}{https://orcid.org/0000-0002-1825-0097}{(Optional) author-specific funding acknowledgements}%TODO mandatory, please use full name; only 1 author per \author macro; first two parameters are mandatory, other parameters can be empty. Please provide at least the name of the affiliation and the country. The full address is optional

\authorrunning{J.\,Q. Public}%TODO mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et al.'

\Copyright{John Q. Public}%TODO mandatory, please use full first names. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/

\ccsdesc[300]{Software and its engineering~Software notations and tools}
\ccsdesc[100]{Software and its engineering~General programming languages}%TODO mandatory: Please choose ACM 2012 classifications from https://dl.acm.org/ccs/ccs_flat.cfm

\keywords{JavaScript, TypeScript, Dynamic Analysis, Declaration Files}%TODO mandatory; please add comma-separated list of keywords

\category{}%optional, e.g. invited paper

\relatedversion{}%optional, e.g. full version hosted on arXiv, HAL, or other repository/website
%\relatedversion{A full version of the paper is available at \url{...}.}

\supplement{}%optional, e.g. related research data, source code, ... hosted on a repository like zenodo, figshare, GitHub, ...

%\funding{(Optional) general funding statement \dots}%optional, to capture a funding statement, which applies to all authors. Please enter author specific funding statements as fifth argument of the \author macro.

% \acknowledgements{I want to thank \dots}%optional

%\nolinenumbers %uncomment to disable line numbering

%\hideLIPIcs  %uncomment to remove references to LIPIcs series (logo, DOI, ...), e.g. when preparing a pre-final version to be uploaded to arXiv or another public repository

%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\EventEditors{John Q. Open and Joan R. Access}
\EventNoEds{2}
\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
\EventShortTitle{CVIT 2016}
\EventAcronym{CVIT}
\EventYear{2016}
\EventDate{December 24--27, 2016}
\EventLocation{Little Whinging, United Kingdom}
\EventLogo{}
\SeriesVolume{42}
\ArticleNo{23}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\maketitle

%TODO mandatory: add short abstract of the document
\begin{abstract}
Developers are starting to write large and complex applications in
TypeScript, a typed dialect of JavaScript. TypeScript applications
integrate JavaScript libraries via typed descriptions of their APIs
called declaration files. DefinitelyTyped is the standard public
repository for these files.
The repository is maintained manually by volunteers, which
is error prone and time consuming. Discrepancies between a
declaration file and the JavaScript implementation lead to
incorrect feedback from the TypeScript IDE and thus to incorrect uses
of the underlying JavaScript library.

This work presents \texttt{dts-generate}, a tool that generates
TypeScript declaration files for JavaScript libraries uploaded to the \NPM{}
registry. It extracts code examples from the documentation written by
the developer, executes the library driven by the examples, gathers
run-time information, and generates a declaration file based on this
information. To evaluate the tool, 244 declaration files were generated and 
compared with the declaration file provided on DefinitelyTyped.
33 files out of 244 had no differences.
% but what about the remaining generated files
\end{abstract}

\section{Introduction}
\label{sec:introduction}
JavaScript has become the most popular language for writing web
applications \cite{github-statistics}. It is also gaining popularity
for back-end applications running in \NodeJS{}, a JavaScript-based
server-side platform. JavaScript is appealing to developers because
its forgiving dynamic typing enables 
them to create simple pieces of code very quickly and proceed on a
trial-and-error basis.

JavaScript was never intended to be more than a
scripting language. Hence, it lacks features for maintaining and evolving large
codebases.
Presently, however, JavaScript is being used for creating large and complex
applications. 
Mistakes such as mistyped property
names and misunderstood or unexpected type coercions cause developers
to spend a significant amount of time in debugging. There is ample
evidence for such mishaps. For example, a 
JavaScript code blog\footnote{\url{https://wtfjs.com}} collects experiences
from developers facing unexpected situations while programming in
JavaScript. \coderef{code:introduction-javascript-wtfs} exposes some
of these unintuitive JavaScript behaviors.

\begin{lstlisting}[
    caption={\textbf{Unexpected JavaScript behavior} - falsy values, \lstinline!typeof!, \lstinline!null! and \lstinline!undefined! operators and type coercion},
    language=JavaScript,
	label=code:introduction-javascript-wtfs,
    float,
    abovecaptionskip=-\medskipamount
]
"0" == false; // true
true == "1.00"; // true
false == "    \n\r\t     "; // true
false == []; // true
0 == []; // true
null == undefined; // true
[1] + 1; //'11'
[2] == "2"; // true
null + undefined + [1, 2, 3] // 'NaN1,2,3'
"hello world".lenth + 1 // NaN | note 'lenth' instead of 'length'
[1, 15, 20, 100].sort() // [ 1, 100, 15, 20 ]
typeof null; // object
null instanceof Object; // false
\end{lstlisting}
% pjt: why unexpected?
% "01" < "00100"; // false

The cognitive load produced by such dynamic typing (which includes
unchecked property names) is not present in other languages that use
build tools based on type information. This insight motivated the
creation of TypeScript, a superset of JavaScript with type
annotations \cite{typescript}. It has become a widely used alternative
among JavaScript developers, because it incorporates features that are
helpful for developing and maintaining large applications
\cite{DBLP:conf/icse/GaoBB17}. TypeScript enables the early detection
of several kinds of run-time errors and the integration of code intelligence
tools like auto-completion in an IDE.

Existing JavaScript libraries can be used in a TypeScript project by
adding a declaration file that contains a description of the library's
API in terms of types. Lots of declaration files for popular
JavaScript libraries have been created in a community effort in the
DefinitelyTyped repository~\cite{definitely-typed-repository}.
At the time of writing this repository contains declaration files for
more than 6000 JavaScript libraries. Unfortunately, most declaration
files in this libraries have been manually created and maintained,
which is error prone and time consuming. As TypeScript takes a
declaration file at face value, discrepancies are not detected at
compile time and the inaccurate code-intelligence features are misleading the
programmer. Moreover, TypeScript does not perform any run-time 
checking on types, either, so that a discrepancy  between the declaration file
and its corresponding JavaScript library can lead to unexpected 
behavior and crashes. Such behavior can lead to developer
frustration, longish debugging sessions, and decreasing confidence in
the tool chain.

Some previous work tackled the problem of improving the quality of
declaration files. Feldthaus et al
\cite{DBLP:conf/oopsla/FeldthausM14} search automatically for
mismatches between a declaration file and implementation code. TSTest
\cite{DBLP:journals/pacmpl/KristensenM17}
adapted feedback-directed random testing to detect discrepancies between a
declaration file and a JavaScript library. Tools like TSInfer and
TSEvolve~\cite{DBLP:conf/fase/KristensenM17} are designed for
assisting the creation of new declaration 
files and supporting the evolution the declaration file when the
corresponding JavaScript library gets modified, respectively. They
rely on an existing static analyzer for JavaScript \cite{TAJS}.
TypeScript itself
developed \texttt{dts-gen}, a tool that generates a
declaration file that is meant to be used as a \emph{starting point
for writing a high-quality declaration file} \cite{dts-gen}.

In this work we present the tool \texttt{dts-generate}
as a first step to explore the possibilities for
generating useful declaration files without the heavy lifting of
static analysis. \texttt{dts-generate} comes with a framework that
supports the generation of declaration files for an existing
JavaScript library published to the \NPM{} registry. The tool gathers
data flow and type information at run-time to generate a declaration
file based on that information.

The novelty of our tool is twofold:
\begin{enumerate}
\item 
  We do not rely on static analysis, which is hard to implement
  soundly and precisely and which is prone to maintenance problems
  when keeping up with JavaScript's yearly language updates.
\item
  Instead we extract example code from the programmer's library
  documentation and rely on dynamic analysis to extract typed usage
  patterns for the library from the example runs.
\end{enumerate}
% The architecture supports the future incorporation of a Symbolic
% Execution Engine that expands the initial code base using the
% signatures in the declaration file. The iterative process of exploring
% new execution paths will refine the generated declaration file in each
% iteration.

% Finally, we generated declaration files for 244 JavaScript libraries
% and evaluated them against the DefinitelyTyped repository.

The contributions of this paper are as follows.
\begin{itemize}
% \item We introduce an architecture that supports generating TypeScript
%   Declaration Files for a given JavaScript Library using run-time
%   information. The architecture supports the future incorporation of a
%   Symbolic Execution Engine that expands the initial code base using
%   the signatures in the declaration file. The iterative process of
%   exploring new execution paths will refine the generated declaration
%   file in each iteration.
\item A framework that extracts code examples from the
  documentation of an \NPM{} package and collects run-time type
  information from running these examples.

\item The tool \texttt{dts-generate}, a command line
  application that generates a valid TypeScript declaration file for a
  specific \NPM{} package using run-time information.

\item A comparator for TypeScript declaration files. This tool is
  necessary for evaluating our framework and also useful to detect
  incompatibilities when evolving JavaScript modules.
\item An evaluation of our framework. We examined all 6029 entries in
  the DefinitelyTyped repository and found 244 sufficiently
  well-documented \NPM{} packages, on which we ran \texttt{dts-generate}
  and compared the outcome with the respective declaration file from the
  DefinitelyTyped repository.

\end{itemize}

\section{Motivating Example}
\label{sec:motivating-example}
The \NPM{} package \texttt{route-parser} is a simple route parsing,
matching, and reversing library for
Javascript\footnote{https://www.npmjs.com/package/route-parser}. It
has about 35,000 weekly downloads and 221 \NPM{} packages depend on it. If
a developer creates or extends TypeScript code that depends on the
\texttt{route-parser} library, the TypeScript
compiler and IDE requires a declaration file for that library to
perform static checking and code completion, respectively. We 
use \texttt{dts-generate} to automatically generate a TypeScript
declaration file for \texttt{route-parser}. The tool downloads the \NPM{} 
package, runs the examples extracted from its documentation, gathers
run-time information, and generates a TypeScript declaration
file. The result is 
shown in \figref{fig:motivating-example-route-parser-vscode} and it is
ready for use in a TypeScript project. For example,  Visual Studio's
code completion runs properly with it
(\figref{fig:motivating-example-route-parser-vscode}). If the
\texttt{route-parser} package gets modified in the future, a new declaration
file can be generated automatically using
\texttt{dts-generate}. Our comparator tool can compare the new file
for incompatibilities with the previous declaration file.

\begin{figure}[tp]
  \centering
  \begin{subfigure}{0.70\linewidth}
    \begin{lstlisting}[language=TypeScript]
export = Route;

declare class Route {
  constructor(spec: string);
  match(path: string): object;
  reverse(params: object): string;
}

declare namespace Route {
}
    \end{lstlisting}
  \end{subfigure}

  \hspace{10.\textwidth}

  \begin{subfigure}{1.\linewidth}
    \centering
    \includegraphics[width=0.7\linewidth]{motivating-example-route-parser-vscode.png}
  \end{subfigure}

  \caption{\textbf{Declaration file for \texttt{route-parser} generated with \texttt{dts-generate}} - Constructor and methods are correctly identified. Declaration file can be correctly used in Visual Studio Code\footnote{https://code.visualstudio.com}.}
  \label{fig:motivating-example-route-parser-vscode}
\end{figure}

In the next section, we examine each step in the generation process in
detail and refer back to this example for concreteness.

\section{The Generation of TypeScript Declaration Files}
\label{sec:gener-typescr-decl}
This section gives an overview of our approach to generating
TypeScript declaration files from JavaScript libraries packaged in
\NPM. \figref{fig:tsd_generation_method_block_diagram} gives a rough
picture of the inner working of our tool \texttt{dts-generate}. The
input is an \NPM{} package and the output is a TypeScript declaration
file for the package if it is ``sufficiently documented'', which we
substantiate in the next subsection.

% We introduce \texttt{dts-generate}, a command line tool which
% generates a valid TypeScript Declaration File for a specific
% JavaScript Library uploaded to the \NPM{} Registry, as explained in
% \figref{fig:tsd_generation_method_block_diagram}. The tool is intended
% to be used on existing, published \NPM{} packages. The generated
% output TypeScript declaration file is a valid file which can be used
% for development and uploaded to the DefinitelyTyped Repository.

\begin{figure}[tp]
    \centering
    \includegraphics[width=1\linewidth]{dts-generate-block-diagram.pdf}
    \caption[dts-generate - Architecture overview]{\textbf{dts-generate - Architecture overview} - Initial code base is retrieved from the \NPM{} package's repository. A valid TypeScript Declaration File is generated using run-time information. A Symbolic Execution Engine creates test cases based on the generated Declaration File and via a feedback loop enriches the code base until the stopping criteria is reached. The final TypeScript Declaration File gets returned. Feedback loop through the Symbolic Execution Engine was not implemented. It can be added in a future to the existing architecture, without modifying the working blocks.}
    \label{fig:tsd_generation_method_block_diagram}
  \end{figure}

As \texttt{dts-generate} is based on run-time information, 
exemplary code fragments that execute the JavaScript library are
needed to obtain 
significant run-time information from running the instrumented library
code.

The examples and the code base of the library are instrumented with
Jalangi \cite{DBLP:conf/sigsoft/SenKBG13} to gather data flow
information and type information at runtime. Jalangi is a configurable
framework for dynamic analysis of JavaScript. It provides several
analysis modules that were extended as needed to retrieve the required
run-time information, which is then saved in a JSON file.

A second independent block uses the run-time information to generate a
TypeScript declaration file. It infers the overall structure of the JavaScript
library, its interfaces, and the types from the run-time information. 
The resulting declaration file is ready for use in the development
process. Its content mimics the usage of the library in the example
code fragments and matches the structure of the
JavaScript library under analysis, so that the JavaScript code
generated after compiling the TypeScript code runs without
interface-related errors.

The command line interface is inspired by the \texttt{dts-gen} package
\cite{dts-gen}. \coderef{code:dts-generate-example} shows
that invoking the package is very simple and the only required
argument is the name of the module published to the \NPM{} registry.

\begin{lstlisting}[
    caption={\textbf{dts-generate usage} - Example of how to generate a declaration file for module \texttt{abs}.},
    language=bash,
	label=code:dts-generate-example,
    float,
    abovecaptionskip=-\medskipamount
]
$ ./dts-generate abs
$ cat output/abs/index.d.ts 
export = Abs;

declare function Abs(input: string): string;
\end{lstlisting}

\subsection{Initial Code Base}
\label{sec:initial-code-base}

To extract run-time information from a JavaScript library, it is
necessary, by definition, to actually execute the code, because the
analysis modules provided by Jalangi to gather information are only
triggered if the instrumented code gets executed.

There are several options to obtain code fragments that drive the
library code:
\begin{enumerate}
\item\label{item:1} execute code that imports the library;
\item\label{item:2} execute the test cases that come with the library;
\item\label{item:3} execute code fragments extracted from the library documentation.
\end{enumerate}

Option~\ref{item:1} does not solve our problem, it just delegates it
to the importing library, which also needs code to drive it. Moreover,
it is costly to download and instrument another package.

We considered option~\ref{item:2} under the assumption that most
libraries would come with test cases. However, there is no standard
for testing JavaScript code so that test cases were difficult to reap
from the \NPM{} packages: they used different directory structures,
different (or no) testing tools, or some packages did not have tests
at all.

In the end, option~\ref{item:3} was the most viable even though there
is no standard for documentation, either. However, almost all repositories
contain README files where the library authors briefly describe what
the code does, which problem it solves, how to install the
application, how to build the code, etc. It is very common that
developers provide code examples in the readme files to show how the
library works and how to use it. This observation holds in particular
for \NPM{} packages, which are generally created to solve a specific
problem of JavaScript development.

Obtaining code examples for a specific \NPM{} package is done in three steps.
\begin{itemize}
\item Obtain the repository url from the package: The command
  \texttt{npm view <PACKAGE> repository.url} can be used for
  retrieving the url of the package's repository.

\item Retrieve the README file from the repository.

\item Extract the code examples from the README file: README files are
  written using Markdown\footnote{\url{https://www.markdownguide.org}}, a
  popular markup language. In a README file, it is customary to write
  code examples within code blocks labeled with the programming
  language, so that syntax highlighting is done correctly. Hence, we
  retrieve the code examples from code blocks labeled \texttt{js} or
  \texttt{javascript}, which both stand for JavaScript in
  Markdown.
\end{itemize}

For example, in the case of \texttt{route-parser} ...

Obtaining the code fragments from the examples provided in the
README files of the repository proved to be an
appropriate and pragmatic way of extracting the developer's
intention and providing an useful initial code base with meaningful
examples, thus avoiding possible cold start problems.

\subsection{Run-time Information Gathering}
The Runtime Information block described in
\figref{fig:tsd_generation_method_block_diagram} gathers
information such as:

\begin{itemize}
  \item Function \lstinline{f} was invoked where parameter
    \lstinline{a} held a value of type \lstinline{string} and
    \lstinline{b} a value of type \lstinline{number}. 
  \item Property \lstinline{foo} of parameter \lstinline{a} of
    function \lstinline{f} was accessed within the function. 
  \item Parameter \lstinline{a} of function \lstinline{f} was used as
    operand for operator \lstinline{==}. 
\end{itemize}

The dynamic analysis framework used for gathering this kind of
information is Jalangi. The configurable analysis modules enable
programming custom callbacks that get triggered with virtually any
JavaScript event. Our instrumentation observes the following events: 
\begin{itemize}
  \item Binary operations, like \lstinline{==}, \lstinline{+} or
    \lstinline{===}.  
  \item Variable declaration.
  \item Function, method, or constructor invocation.
  \item Access to an object's property.
  \item Unary operations, like \lstinline{!} or \lstinline{typeof}.
\end{itemize}

The implementation stores these observations as entities called
\texttt{interactions}. They are used for translating, modifying, and
aggregating Jalangi's raw event information to get an application
specific data representation. The run-time information is saved as a
JSON file that can be used for later processing. The tool is written
in JavaScript and runs in \NodeJS{} in a Docker container.

\subsection{TypeScript Declaration File Generation}
\label{sec:typescr-decl-file}
\paragraph*{Overview}
The actual generation of the declaration file is the next step in the
pipeline after gathering the run-time information, as shown in
\figref{fig:tsd_generation_method_block_diagram}. It is a lightweight,
simple and fast application, which does not interact with the actual
JavaScript module at run-time. Instead, it uses the JSON output file
containing the run-time information and generates a TypeScript
declaration file which is use ready to be used in a TypeScript
project. The tool itself is written in TypeScript and runs in a
Docker container in \NodeJS.

\paragraph*{Templates}
TypeScript provides templates for writing declaration files\footnote{https://www.typescriptlang.org/docs/handbook/declaration-files/templates.html} and each template corresponds to a different way of exporting a JavaScript module. The tool uses different fields from the run-time information to detect how the module is being used in order to choose the right templates accordingly. The implemented templates are \lstinline{module}, \lstinline{module-class} and \lstinline{module-function}.

\paragraph*{Interfaces}
Finally, interfaces are created by exploring \lstinline{getField} and \lstinline{methodCall} interactions from the run-time information. The code will gather the interactions for a specific argument and build the interface by incrementally adding new properties. Interactions within the \lstinline{followingInteractions} field are recursively traversed, building a new interface in each recursion level.

\section{Evaluation}
\label{sec:dts-generate-evaluation}
After generating a declaration file for a published \NPM{} package, it is necessary to evaluate the quality of it. It was decided to compare the generated declaration file against the one uploaded to the DefinitelyTyped repository for the same module, as shown in \figref{fig:evaluation-diagram}.

\begin{figure}[tp]
    \begin{centering}
        {\includegraphics[width=1\textwidth]{evaluation-diagram.pdf}}
        \caption[Evaluation against DefinitelyTyped Repository]{\textbf{Evaluation of generated declaration files against DefinitelyTyped Repository} - A parser transforms the generated declaration file and the equivalent file in the DefinitelyTyped repository into a JSON file using the TypeScript Compiler API \cite{typescript-compiler-api}. Comparison is then performed on the JSON files, i.e. not on the declaration files.}
        \label{fig:evaluation-diagram}
    \end{centering}
\end{figure}

\paragraph*{Parsing}
Before comparison, declaration files need to be parsed so that both of them share the same structure. It was achieved by using the TypeScript Compiler API, a library developed by Microsoft that allows to traverse the Abstract Syntax Tree in an easy and intuitive way \cite{typescript-compiler-api}.

The parsing consists in creating a structure where declared \lstinline{interfaces}, \lstinline{functions}, \lstinline{classes} and \lstinline{namespaces} are stored separately. Function arguments are correctly described, identifying complex types like union types or callbacks. Optional parameters are also identified. For \lstinline{classes}, a distinction between the constructor and methods is made. Finally, syntax and semantic errors are also checked by the TypeScript Compiler API.

The tool is called \texttt{parse-dts} and is naturally written in TypeScript. It also runs in \NodeJS{} in its corresponding Docker container.

\paragraph*{Comparator}
An independent tool will compare two parsed declaration files. As described in \figref{fig:evaluation-diagram}, the comparator will compare the generated declaration file against the corresponding file in the DefinitelyTyped repository.

It was discovered that the implementation was easier when focusing on each template independently. For this implementation, only the \lstinline{module-function} was considered for the comparison.

The following criteria were applied:
\begin{itemize}
  \item Number of declared functions: Checks the number of declared and exported functions for each of both declaration files.
  \item Name of declared functions: Checks whether both of the declaration files declared a function with the same name.
  \item Number of parameters: Checks the number of parameters of the declared functions. This is checked for optional and non-optional parameters.
  \item Interfaces: Checks the number of declared interfaces and the fields within those interfaces.
  \item Errors: Indicates whether there are errors in the declaration files.
\end{itemize}

\section{Results}
\label{sec:results}

\begin{quotation}
  The conducted experiments included tests that consisted of replacing
  a specific type definition from DefinitelyTyped
  \cite{definitely-typed-repository} with the one generated in the
  experiments: TypeScript compilation was successful, the generated
  JavaScript code ran without errors and code intelligence features
  performed by IDEs like code completion worked as expected.
\end{quotation}

Declaration files were generated for existing modules uploaded to the \NPM{} registry. The DefinitelyTyped repository was used as a benchmark. Each one of the generated files was compared against the corresponding declaration file already uploaded to the repository.

Figure \ref{fig:experiments-overall-funnel} shows that a declaration file was generated for 244 modules out of 6029 modules. Samples of the generated declaration files for templates \lstinline{module}, \lstinline{module-class} and \lstinline{module-function} are presented in \secref{sec:experiments-declaration-files-generation}.

\definecolor{myellow}{RGB}{228,212,0}
\definecolor{mgreen}{RGB}{5,104,57}

\newcommand\funnel[3]{%
\pgfmathsetmacro\mwid{(0.3+\val*0.06)}
\pgfmathsetmacro\mradius{(\val*0.01 + 1)}
\pgfmathsetmacro\mheight{(\val*0.003 + 0.4)}
\pgfmathsetmacro\marc{\mwid-.4}
    \begin{scope}[%
        shift={(0,#1)}, 
        line width=.05pt, 
        %x=5mm, 
        scale=0.7,
        yshift=\xi*0.05
        ]
    \draw[black,bottom color=#2, top color=#2] (-\mwid,0) -- (-\mwid+.4,-\mheight) arc (190:350:\marc cm and \mradius mm) -- (\mwid,0);
    \draw[black,fill=#3] (0,0) ellipse (\mwid cm and \mradius mm);
    \path (-\mwid,0) -- (-\mwid+.4,-\mheight) coordinate[midway] (a\xi);
    \end{scope}
}

\begin{figure}[tp]
	\hspace*{-0.16\textwidth}
	\centering
	\begin{tikzpicture}
		\foreach \val
				[%
				count=\xi starting from 0, 
				evaluate=\xi as \shadecolor using int(25*\xi),
				evaluate=\xi as \coord using int(\xi-12)
				]
			in {
				4.05,
				7.33,
				24.02,
				37.49,
				71.19,
				82.50,
				100.00
			}{
				\funnel{\coord}{mgreen!\shadecolor !myellow}{mgreen!\shadecolor !myellow}
			}

		\node[left=0.02\textwidth of a0] {Generated Declaration Files};
		\node[right=0.04\textwidth of a0] {\textbf{244}};

		\node[left=0.02\textwidth of a1] {Run-time Information};
		\node[right=0.06\textwidth of a1] {\textbf{442}};

		\node[left=0.02\textwidth of a2] {Working Examples};
		\node[right=0.17\textwidth of a2] {\textbf{1448}};

		\node[left=0.02\textwidth of a3] {Code Examples};
		\node[right=0.25\textwidth of a3] {\textbf{2260}};

		\node[left=0.02\textwidth of a4] {README file};
		\node[right=0.45\textwidth of a4] {\textbf{4292}};

		\node[left=0.02\textwidth of a5] {Github Repository};
		\node[right=0.52\textwidth of a5] {\textbf{4974}};

		\node[align=right,left=0.02\textwidth of a6, text width=0.25\textwidth] {Definitely Typed Modules};
		\node[right=0.62\textwidth of a6] {\textbf{6029}};

	\end{tikzpicture}
	\caption[Number of analyzed modules for each stage of the experiment]{\textbf{Number of analyzed modules for each stage of the experiment} - A TypeScript Declaration File was generated for only 244 modules, out of 6029 modules in the DefinitelyTyped Repository. It was possible to gather valid run-time information for only 25\% of the modules for which a Code Example was extracted.}
	\label{fig:experiments-overall-funnel}
\end{figure}

\subsection{Code Examples}
Retrieving the code examples from the JavaScript libraries' repositories proved to be a pragmatic way of capturing the types. However, as shown in \figref{fig:experiments-overall-funnel}, working code examples for only 2260 modules could be retrieved. The process of getting a valid code example for a module is divided in 4 blocks:
\begin{itemize}
	\item Extracting repositories url.
	\item Extracting readme files.
	\item Extracting code examples from README files.
	\item Executing code examples and discarding failing ones.
\end{itemize}

The results obtained for each on of them are described in the following sections.

\paragraph*{Repositories URL}
The url of the repositories could be retrieved for only 4974 modules. More than 1000 modules do not have the repository entry in their corresponding \texttt{package.json} files. Therefore, the \texttt{npm view <module> repository.url} command returns an empty value. This is even happening for important modules like \texttt{ace}.

\paragraph*{Readme Files}
700 modules simply do not have a readme file in their repositories. The implementation does contemplate, however, different naming conventions like \texttt{readme.md} or \texttt{README.md}.

\paragraph*{Code Examples Extraction}
The 50\% loss is mainly explained because developers did not wrap their code around a block using the \texttt{javascript} or \texttt{js} tags. Counting with code examples for 2200 modules was considered to be enough for evaluating the generation of declaration files.

\paragraph*{Code Examples Execution}
The 2260 extracted code examples were executed by installing the required packages and running the code as a \NodeJS{} application. Working and functional code examples could only be extracted for 1448 modules. 812 modules did not run correctly and were discarded. Some failing samples were analyzed and there were mainly two reasons for the failure:
\begin{enumerate}
	\item The code example had been properly extracted but the code itself was not working. It was executing the library in an unsupported way, hence the error at run-time.
	\item The extracted code example was not intended to be executed or it was not even valid JavaScript code.
\end{enumerate}

\subsection{Declaration Files Generation} \label{sec:experiments-declaration-files-generation}

The following section exhibits some samples of the 244 generated declaration files. It shows some results for each of the implemented templates: \lstinline{module}, \lstinline{module-function} and \lstinline{module-class}.

\figref{fig:experiments-results-module-function} shows the generated declaration files for simple modules like \texttt{abs}, \texttt{dirname-regex} and \texttt{escape-html}. All of them were generated using the \texttt{module-function} template. There are no differences between the generated files and the corresponding declaration files uploaded to DefinitelyTyped.

\begin{figure}[tp]
    \centering
    \begin{subfigure}{0.48\linewidth}
      \begin{lstlisting}[language=TypeScript]
export = Abs;

declare function Abs(input: string): string;
      \end{lstlisting}
      \caption{abs/index.d.ts - Generated}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.48\linewidth}
      \begin{lstlisting}[language=TypeScript]
declare function Abs(input: string): string;
export = Abs;
      \end{lstlisting}
      \caption{abs/index.d.ts - DefinitelyTyped}
    \end{subfigure}

    \begin{subfigure}{0.48\linewidth}
        \begin{lstlisting}[language=TypeScript]
export = DirnameRegex;

declare function DirnameRegex(): RegExp;
        \end{lstlisting}
        \caption{dirname-regex/index.d.ts - Generated}
      \end{subfigure}
      \hfill
      \begin{subfigure}{0.48\linewidth}
        \begin{lstlisting}[language=TypeScript]
export = dirnameRegex;

declare function dirnameRegex(): RegExp;
        \end{lstlisting}
        \caption{dirname-regex/index.d.ts - DefinitelyTyped}
      \end{subfigure}

      \begin{subfigure}{0.48\linewidth}
        \begin{lstlisting}[language=TypeScript]
export = EscapeHtml;

declare function EscapeHtml(string: string): string;
        \end{lstlisting}
        \caption{escape-html/index.d.ts - Generated}
      \end{subfigure}
      \hfill
      \begin{subfigure}{0.48\linewidth}
        \begin{lstlisting}[language=TypeScript]
declare function escapeHTML(text: string): string;
declare namespace escapeHTML { }

export = escapeHTML;
        \end{lstlisting}
        \caption{escape-html/index.d.ts - DefinitelyTyped}
        \label{fig:experiments-results-module-function}
      \end{subfigure}

    \caption{\textbf{Module-function results} - Results are shown for modules \texttt{abs}, \texttt{dirname-regex}, \texttt{escape-html}. On the left side the generated declaration file with \lstinline{dts-generate}. On the right side the corresponding file in the DefinitelyTyped repository. Functions are correctly detected and input types are accurately inferred. Both files are parsed for comparison, as explained in \secref{sec:dts-generate-evaluation}. Therefore, subtle differences in the syntax between both files are not important.}
\end{figure}

Templates of type \lstinline{module-class} are shown for modules \texttt{flake-idgen}, \texttt{route-parser} and \texttt{timer-machine} in \figref{fig:experiments-results-module-class-flake-idgen} and \figref{fig:experiments-results-module-class-timer-machine}, respectively. Properties of interfaces and class methods are correctly generated. Optional parameters are not detected, as it was not considered for the implementation. Finally, \lstinline{module} template is presented for \texttt{is-uuid} module in \figref{fig:experiments-results-module-is-uuid}.

\begin{figure}[tp]
    \centering
    \begin{subfigure}{0.48\linewidth}
      \begin{lstlisting}[language=TypeScript]
export = FlakeIdgen;

declare class FlakeIdgen {
    constructor(options: FlakeIdgen.I__options);
    next(cb: undefined): Buffer;
}

declare namespace FlakeIdgen {
    export interface I__options {
        'id': undefined;
        'datacenter': number;
        'worker': number;
        'epoch': undefined;
        'seqMask': undefined;
    }
}
      \end{lstlisting}
      \caption{flake-idgen/index.d.ts - Generated}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.48\linewidth}
      \begin{lstlisting}[language=TypeScript]
interface ConstructorOptions {
    datacenter?: number;
    worker?: number;
    id?: number;
    epoch?: number;
    seqMask?: number;
}

declare namespace FlakeId { }

declare class FlakeId {
    constructor(options?: ConstructorOptions);
    next(callback?: (err: Error, id: Buffer) => void): Buffer;
}

export = FlakeId;        
      \end{lstlisting}
      \caption{flake-idgen/index.d.ts - DefinitelyTyped}
    \end{subfigure}

    \caption{\textbf{Module-class results | flake-idgen} - Parameters of interface \texttt{ConstructorOptions} are correctly detected. Name of interface differs since it is automatically generated based on the name of the argument variable. Optional properties were not implemented, hence the \lstinline{undefined} type for some properties. Analogously, callback \texttt{cb} is inferred as \lstinline{undefined}.}
    \label{fig:experiments-results-module-class-flake-idgen}
\end{figure}

\begin{figure}[tp]
    \centering
    \begin{subfigure}{0.48\linewidth}
      \begin{lstlisting}[language=TypeScript]
export = Timer;

declare class Timer {
    constructor(start: undefined);
    start(): boolean;
    isStopped(): boolean;
    emit(): boolean;
    stop(): boolean;
    isStarted(): boolean;
    timeFromStart(): number;
    time(): number;
}

declare namespace Timer {
}
      \end{lstlisting}
      \caption{timer-machine/index.d.ts - Generated}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.48\linewidth}
      \begin{lstlisting}[language=TypeScript]
export as namespace Timer;
export = Timer;

declare namespace Timer {
    type TimerEvent = "start" | "stop" | "time";
}

declare class Timer {
    static get(reference: string): Timer;
    static destroy(reference: string): Timer;

    constructor(started?: boolean);

    isStarted(): boolean;
    isStopped(): boolean;
    start(): void;
    timeFromStart(): number;
    stop(): void;
    time(): number;
    toggle(): void;
    emitTime(): void;
    valueOf(): number;
    on(event: Timer.TimerEvent, callback?: () => void): void;
}
      \end{lstlisting}
      \caption{timer-machine/index.d.ts - DefinitelyTyped}
    \end{subfigure}

    \caption{\textbf{Module-class results | timer-machine} - Parameter \lstinline{started} is inferred as \lstinline{undefined} instead of marking it as optional. Methods that were not executed do not appear in the generated declaration file.}
    \label{fig:experiments-results-module-class-timer-machine}
\end{figure}

\begin{figure}[tp]
    \centering
    \begin{subfigure}{0.48\linewidth}
      \begin{lstlisting}[language=TypeScript]
export function v1(str: string): boolean;
export function v2(str: string): boolean;
export function v3(str: string): boolean;
export function v4(str: string): boolean;
export function v5(str: string): boolean;
      \end{lstlisting}
      \caption{is-uuid/index.d.ts - Generated}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.48\linewidth}
      \begin{lstlisting}[language=TypeScript]
export function v1(value: string): boolean;
export function v2(value: string): boolean;
export function v3(value: string): boolean;
export function v4(value: string): boolean;
export function v5(value: string): boolean;
export function nil(value: string): boolean;
export function anyNonNil(value: string): boolean;
      \end{lstlisting}
      \caption{is-uuid/index.d.ts - DefinitelyTyped}
    \end{subfigure}

    \caption{\textbf{Module results | is-uuid} - Methods that were not executed are not included in the declaration file.}
    \label{fig:experiments-results-module-is-uuid}
\end{figure}

It is worth mentioning that for some libraries the declaration file in DefinitelyTyped was not correct. For example, for \texttt{datadog-metrics}, some properties of an interface were included in the generated declaration file but they were not present in the one in DefinitelyTyped. However, as shown in \figref{fig:experiments-results-module-datadog-metrics}, the properties are indeed used in the source code and should be included.

\begin{figure}[tp]
    \centering
    \begin{subfigure}{0.48\linewidth}
      \begin{lstlisting}[language=TypeScript]
export interface I__opts {
    'aggregator': undefined;
    'defaultTags': Array<any>;
    'reporter': undefined;
    'apiKey': string;
    'appKey': undefined;
    'agent': undefined;
    'host': string;
    'prefix': string;
    'flushIntervalSeconds': number;
}

export class BufferedMetricsLogger {
    constructor(opts: I__opts);
    // ...
}
      \end{lstlisting}
      \caption{datadog-metrics/index.d.ts - Generated}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.48\linewidth}
      \begin{lstlisting}[language=TypeScript]
export interface LoggerOptions {
    apiKey?: string;
    appKey?: string;
    defaultTags?: string[];
    flushIntervalSeconds?: number;
    host?: string;
    prefix?: string;
}

export class BufferedMetricsLogger {
    constructor(
        options: LoggerOptions
    );
    // ...
}
      \end{lstlisting}
      \caption{datadog-metrics/index.d.ts - DefinitelyTyped}
    \end{subfigure}

    \begin{subfigure}{0.80\linewidth}
        \begin{lstlisting}[language=TypeScript]
function BufferedMetricsLogger(opts) {
    this.aggregator = opts.aggregator || new Aggregator(opts.defaultTags);
    this.reporter = opts.reporter || new DataDogReporter(opts.apiKey, opts.appKey, opts.agent);
    this.host = opts.host;
    this.prefix = opts.prefix || '';
    this.flushIntervalSeconds = opts.flushIntervalSeconds;

    // ...
    // ...
}
        \end{lstlisting}
        \caption{datadog-metrics/logger.js}
        \label{fig:subfloat-datadog-logger}
      \end{subfigure}

    \caption{\textbf{Missing properties | datadog-metrics} - Properties \texttt{aggregator} and \texttt{reporter} are not in the DefinitelyTyped version, but they appear in the generated declaration file. However, they are indeed used by the library, as exposed in lines 2 and 3 of the library's source code shown in \protect\subref{fig:subfloat-datadog-logger}.}
    \label{fig:experiments-results-module-datadog-metrics}
\end{figure}

\subsection{Evaluation}
As shown in \figref{fig:experiments-typescript-templates-distribution-definitely-typed}, 20\% of the declaration files in DefinitelyTyped are written using the \lstinline{module-function}. However, 57\% of the 244 generated declaration files are written with the \lstinline{module-function} template. Additionally, the complexity of evaluating declaration files written with the \lstinline{module-function} is considerably lower than for other templates. The evaluation for templates \lstinline{module-class} and \lstinline{module} was not implemented.

33 out of 116 evaluated modules have no difference with their corresponding declaration file in DefinitelyTyped.

\begin{figure}[tp]
	\centering
	\begin{tikzpicture}
		\begin{axis}[
			ybar,
            width=0.7\textwidth,
            scale=0.8,
			ybar=0pt,
			ymax=100,
			enlargelimits=0.15,
			bar width=0.04\textwidth,
			legend columns=1,
			legend cell align={left},
			legend style={
                at={(0.6,0.95)},
                anchor=north,
                nodes={scale=0.8, transform shape}
            },
			symbolic x coords={module-class,module-function,module},
			xtick=data,
			ylabel=Percentage,
			nodes near coords, 
			nodes near coords align={vertical}
		]
		\addplot coordinates {
			(module-class, 8)
			(module-function, 20)
			(module, 72)
		};

		\addplot coordinates {
			(module-class, 12)
			(module-function, 57)
			(module, 31)
		};
		\legend{DefinitelyTyped (Total: 6029), Generated (Total: 244)}
		\end{axis}
	\end{tikzpicture}

	\caption[TypeScript templates distribution | Generated \& DefinitelyTyped]{\textbf{TypeScript templates distribution | Generated \& DefinitelyTyped} - Out of a total of 6029, 72\% of the modules uploaded to the DefinitelyTyped repository use the \lstinline{module} template and only 20\% use the \lstinline{module-function} one. However, 57\% of the 244 generated declaration files use the \lstinline{module-function} template.
	}

	\label{fig:experiments-typescript-templates-distribution-definitely-typed}
\end{figure}

\section{Related Work}
\label{sec:related-work}
\paragraph*{Microsoft's dts-gen}
Microsoft developed \texttt{dts-gen}, a tool that creates starter declaration files for JavaScript libraries \cite{dts-gen}. Its documentation states that the result is however intended to be only used a starting point. The outcome needs to be refined afterwards by the developers.

The tool analyzes the shape of the objects at runtime after initialization without executing the library. This results in many variables being inferred as \lstinline[language={}]{any}. \coderef{code:related-work-dts-gen-example} shows an example for module \lstinline[language={}]{abs}.

The solution presented in this work, however, is intended to generate declaration files that are ready to be uploaded to DefinitelyTyped without further manual intervention. Any amount of manual work that a developer needs to do on a declaration file after updating JavaScript code increases the risk for having discrepancies between the declaration file and the implementation.

Formal aspects like applying the right template and using the correct syntax are perfectly covered by \texttt{dts-gen}.

\begin{lstlisting}[
    language=bash,
    caption={Microsoft's dts-gen example - A declaration file for module \lstinline!abs! is generated. Types are inferred as \lstinline!any!. The correct \lstinline!module-function! template is used.},
	label=code:related-work-dts-gen-example,
    float,
    abovecaptionskip=-\medskipamount
]
$ npm i -g dts-gen
$ npm i -g abs
$ dts-gen -m abs
Wrote 5 lines to abs.d.ts.

$ cat abs.d.ts
/** Declaration file generated by dts-gen */

export = abs;

declare function abs(input: any): any;
\end{lstlisting}

\paragraph*{TSInfer \& TSEvolve}
TSInfer and TSEvolve are presented as part of TSTools \cite{DBLP:conf/fase/KristensenM17}. Both tools are the continuation of TSCheck \cite{DBLP:conf/oopsla/FeldthausM14}, a tool for looking for mismatches between a declaration file and an implementation.

TSInfer proceeds in a similar way than TSCheck. It initializes the library in a browser and it records a snapshot of the resulting state and then it performs a light weight static analysis on all the functions and objects stored in the snapshot.

The abstraction and the constraints they introduced as part of the static analysis tools for inferring the types have room for improvement. A run-time based approach like the one presented in our work will provide more accurate information, thus generating more precise declaration files.

Since they analyze the objects and functions stored in the snapshot, they faced the problem of including in the declaration file internal methods and properties that developers wanted to hide. Run-time information would have informed that the developer has no intention of exposing such methods.

Moreover, TSEvolve performs a differential analysis on the changes made to a JavaScript library in order to determine intentional discrepancies between declaration files of two consecutive versions. We consider that a differential analysis may not be needed. If the developer's intention is accurately extracted and the execution code clearly represents that intention then the generated declaration file would already describe the newer version of a library without the need of a differential analysis.

\paragraph*{TSTest}
TSTest is a tool that checks for mismatches between a declaration file and a JavaScript implementation \cite{DBLP:journals/pacmpl/KristensenM17}. It applies feedback-directed random testing for generating type test scripts. These scripts will execute the library in order to check if it behaves the way it is described in the declaration file. TSTest also provides concrete executions for mismatches.

We evaluated the generated declaration files comparing them to the declaration files uploaded to DefinitelyTyped. The disadvantage of doing this is that since the uploaded files are written manually, they could already contain mismatches with the JavaScript implementation. However, it is a suitable choice for a development stage since it is used as a baseline.

In a final stage, declaration files need to be checked against the proper JavaScript implementation and TSTest has to be definitely taken into account.

\section{Conclusion}
\label{sec:conclusion}
We have presented \texttt{dts-generate}, a tool for generating a TypeScript declaration file for a specific JavaScript library. The tool downloads code samples written by the developers from the library's repository. It uses these samples to execute the library and gather data flow and type information. The tool finally generates a TypeScript declaration file based on the information gathered at run-time.

We developed an architecture that supports the automatic generation of declaration files for specific JavaScript libraries without additional manual tasks. The architecture contemplates a future incorporation of a Symbolic Execution Engine that refines the initial code base enabling the exploration of new execution paths. However not implemented in this work, its incorporation would result in small incremental modifications to the presented architecture as it is considered to only expand the existing code base.

Building an end-to-end solution for the generation of TypeScript declaration files was prioritized over type inference accuracy. Consequently, types were taken over from the values at run-time. Since developers expose through code how a library should be used, obtaining the types from the code examples extracted from the repositories proved to be a pragmatic and effective approximation, enabling to work on specific aspects regarding the TypeScript declaration file generation itself.

We built a mechanism to automatically create declaration files for potentially every module uploaded to DefinitelyTyped. We managed to generate declaration files for 244 modules. We compared the results against the corresponding files uploaded to DefinitelyTyped by creating a TypeScript declaration files parser and a comparator.

We exposed the fundamental aspect of capturing the developer's intention when inferring types in JavaScript. Instead of applying constraints and restrictions for operations with certain types, we presented a proposal where common practices are favored. Uncommon usage is not forbidden but greatly disfavored. Accordingly, we collected evidence regarding the usage of JavaScript operators by analyzing 400 libraries.

Finally, the architecture is composed of different blocks that interact with each other. Each block is independent and has a well defined behavior as well as clear input and output values. As a result, each block can be independently and simultaneously improved.

\bibliography{main}

\end{document}