### 2021-05-26 mplr-2021/programming21-reviews-answered.txt
logged

> ----------------------- REVIEW 1 ---------------------
> SUBMISSION: 28
> TITLE: Generation of TypeScript Declaration Files from JavaScript Code
> AUTHORS: Fernando Cristiani and Peter Thiemann
> 
> ----------- Overall evaluation -----------
> SCORE: -1 (major revision)
> ----- TEXT:
> Automatically generating TypeScript type signatures is a worthwhile
> goal; a mechanical process for generating such declarations would be
> less tedious, less error-prone, and more scalable. Pursuing a run-time
> trace-based technique is also compelling, given the significant
> difficulties that dynamically typed languages pose for static
> analysis. The initial results of the tool are promising: 244
> TypeScript declarations were automatically generated from existing
> module definitions, and 70 of these could be compared and validated
> against manually written declarations.
> 
> However, I believe the paper falls short both in terms of technical
> formulation as well as evidence of practical utility.
> 
> With respect to technical approach, the paper says (p.3/4) that the
> use of examples and dynamic analysis, rather than static analysis, is
> a novelty of the proposed tool. But a similar approach has been
> pursued previously:
> 
> * Trace Typing: An Approach for Evaluating Retrofitted Type Systems (Andreasen et al. ECOOP 2016)
> 
> * Just-in-Time Static Type Checking for Dynamic Languages (Ren et al. PLDI 2016)
> 
> * Dynamic Inference of Static Types for Ruby (An et al. POPL 2011)
> 
> * JSTrace: Run-time Type-Discovery for JavaScript (Saftoiu et al. 2010, > http://cs.brown.edu/research/plt/dl/flowtyping/v1/saftoiu-typediscovery-thesis.pdf)
> 
> How do the proposed techniques compare? Among these, trace typing is
> particularly relevant: like the current paper, it uses Jalangi for
> instrumentation.

Added discussion to the related work section.

> Besides not discussing this related work, the descriptions in Sections
> 4.2 and the end of 4.3 are rather informal: What exactly is the subset
> of JavaScript/TypeScript that is supported?

ECMAScript 5.1

> What is the instrumented
> evaluator computing? How are types --- which include unions and
> intersections (at least for function types) --- being computed from
> the run-time information? The lack of precise details makes it
> difficult to understand what the tool currently does, how plausible it
> might be to extend to support more of JavaScript/TypeScript in the
> future, and how it relates the related work described above.

Added some details about type computation at end of section 4.

> With respect to the implemented tool, it is not clear how close it is
> to being useful in practice. Many TypeScript features, as described in
> Table 1 and in the Evaluation, are not yet supported. For the 244
> modules for which declarations could be generated (and for the 70
> which could be compared to existing manual declarations): How many
> lines of code are in the modules? How many definitions are exported?
> How many examples were provided? How many examples were added to the
> 10 modules that were marked solvable? And how do these characteristics
> compare to the remaining Definitely Typed modules not currently
> supported? Especially because of the questions about technical
> formulation and novelty, it seems important for the system to come
> with a more comprehensive practical evaluation.
> 
> Overall, there is a lot of potential for this work --- especially on
> the practical side, for building a useful system --- but I think the
> results are preliminary to publish in their current form.
> 
> 
> *** Additional Comments and Questions ***
> 
> Fig 1: Does the tool generate declarations like this one? (If so, where
> does the name l__opts come from?)

Added explanation and a little cleanup.

> p.6: I think the reliance on Jalangi, as a black box, should be
> described earlier, along with related work.

Done.

> Fig 2: I spent some time wondering how the module names in (b) and (f)
> were chosen (the former appears to be camel case version of the module
> name, and the latter follows the name in the example program). Some of
> this explanation comes later, but perhaps it's worth adding more
> explanation here as this seems to be the primary place in the paper
> where the different kind of templates are explained.

Said that for Globtoregexp.I__opts already!

> Fig 3: Why not depict the example extraction step --- which sometimes
> includes manual addition of examples --- here? And consider
> representing Jalangi's role in this architecture.

Fernando?
F: Added new images:
  - mplr-2021/graphics/dts-generate-block-diagram-with-code-examples-block.pdf
  - mplr-2021/graphics/dts-generate-block-diagram-with-jalangi.pdf

> Sec 4.2/4.3: As mentioned, I think a lot more detail is warranted
> about the instrumentation and the process for computing types from
> traces.
> 
> "Execution of Code Examples": Interesting to see how many examples
> were perhaps out-of-date or were not proper examples. Were there any
> interesting patterns worth discussing?
> 
> p.17: It's encouraging to see how the proposed tool can fix some
> errors found in manual type declarations!
> 
> "Literal types vs string": If considering asking developers to mark
> strings, instead why not ask them to directly use, e.g. the "Enums" or
> "Discriminated Unions" (https://www.typescriptlang.org/docs/handbook)?

Misconception! They write JavaScript, not TypeScript! Added a
paragraph of clarification.

> "Any": Ditto above; why not ask developers to choose the any type?
> 
> "Further types": Inferring indexed types, dependent types, etc. from
> run-time traces would seem to require quite significant technical
> extensions to the approach described here.
> 
> TSInfer: Is it fundamental that TSInfer apparently generates
> signatures for helpers? Could there not be a simple syntactic analysis
> that identifies top-level definitions, to be include in a declaration
> file?
> 
> 
> *** Minor Comments and Typos ***
> 
> Fig 1: Maybe I'm missing it, but I don't see Footnote 6 anywhere.j

FIXED

> p.5: "globToRegexp" ==> "GlobToRegexp" twice

FIXED

> p.8: I'm not sure why "syntax highlighting is done correctly" is
> noteworthy here.

REMOVED

> Fig 4: Is this referenced in the text?

FIXED

> p.15: "loose" ==> "lose"
>

FIXED

> 
> ----------------------- REVIEW 2 ---------------------
> SUBMISSION: 28
> TITLE: Generation of TypeScript Declaration Files from JavaScript Code
> AUTHORS: Fernando Cristiani and Peter Thiemann
> 
> ----------- Overall evaluation -----------
> SCORE: -1 (major revision)
> ----- TEXT:
> # Summary
> 
> The paper describes the tool dts-generate and its necessary companions to generate TypeScript declaration files for > JavaScript packages.
> These JavaScript packages need to provide a repository with a dedicated README that includes documentation efforts in the > form of exemplary usages.
> Based on the run-time information gathered by running these example, `dts-generate` produces the final TypeScript > declaration file after an additional analysis step that determines the scope of the file at hand (module, module-class, > module-function).
> The tool was evaluated based on all NPM packages that fulfilled the documentation requirement sufficiently and > community-driven declaration files; these declaration files are compared against the generated ones using an additional > tooling machinery developed by the authors. 
> 
> # My two cents
> 
> The general idea of the paper as well as the benefit of this idea (as well as the engineering effort) is worthwhile.
> The writing, on the other hand, does not feel mature enough for an acceptance.
> Unfortunately, I have problems myself to pin the exact reasons for this feeling.
> Let me try.
> 
> For an "empirical perspective" the evaluation (or here: result) section does not has the necessary depth to convince the > reader with, nor does it contain empirical methods (that I'm also not an expert for) that would strengthen the choice of this > perspective.
> Instead the results include further explanations of the engineering efforts necessary for the evaluation (how code examples > were extracted), single example extracts from the generated and original declaration files.
> The section concludes with a rather small paragraph (6.3 Evaluation) that only states that 70 files could be tested and all > qualified for a positive result.
> One deficit of this shallow evaluation origins from the small number of final declaration files that qualified for the evaluation.
> See some of the major and minor comments for more direct questions/remarks relating to specific sentences tackling this > deficiency.
> 
> The preceding part of the paper follows this rather shallow evaluation.
> The description of the engineering effort is given in detail, but not in depth.
> I guess these shallow descriptions are sufficient for a paper with an empirical emphasis, but since the focus of the current > version of paper is not the empirical part, I was wondering if an engineering perspective would be more suitable.
> In its current form, the description of the engineering effort needs additional work to make the shift to such an perspective > (what's the output the run-time information provided by Jalangi that dts-generate makes use of?).
> 
> All in all, I think that the paper needs a major revision (and shift of focus) in order to classify as empirical perspective or > needs to change the perspective to an engineering paper with additional efforts needed to fulfil the reader's wish to get more > in-depth insight to the technical part of this work.
> 
> ## Major comments
> 
> p. 1: I have the feeling that the abstract was not prepared in
> accordance to the specifications of the journal including the
> keywords context, inquiry, approach, knowledge, grounding and importance.
> It's not that I do not get an impression of the work presented by
> the paper, but it just feels like an "ordinary abstract" and
> not one that explicitly tackles the key questions behing the keywords above.
> 
> p. 4: "We examined all 6029 entries in the DefinitelyTyped
> repository and found 244 sufficiently well documented NPM
> packages," - it's a shame that the final portion of repositories that can be tested is less than 5% of the starting value.
> 
> p. 15: "In this step, we loose another 50% of modules! This loss is
> mainly explained because some developers do not wrap
> their code in a block using the javascript or js tags. However, as
> we are still left with code examples for 2260 modules, we did
> not further look into code extraction as this number was considered
> sufficient for evaluating the generation of declaration
> files." - Given the low value for the final number of repositories,
> you should definitely reconsider adapting your extraction
> mechanism to be more robust!
> 
> p. 17: "We were unable to generate a declaration file based on the
> module-class template. All modules corresponding to this template
> exhibited at least one of the unimplemented TypeScript features." -
> I was wondering if could make sense to adapt
> the tooling machinery to leave out only the definitions that consist
> of unimplemented features, instead of leaving out a
> whole module if one of its definition.
> 
> p. 18: "The remaining 174 files contained at least one of the
> unimplemented TypeScript features so that they could not be
> considered." - This sentence sounds like you indeed generate as much
> as possible and filter out only layer --- how does this
> statement relate my previous concern?
> 
> ## Minor comments
> 
> p. 5: "The choice of the template depends on the structure of the
> underlying JavaScript library:..." - here the following
> listing/itemize environment seems to be missing, at least, I think that a itemize would be the better layout to use here.

DONE.

> p. 14: "dts-compare does not contemplate differences in function
> return types, because interface inference of function
> return types is not covered in the current version of dts-generate."
> - Why is it that the inference for return types is that
> different from parameters? You infer the parameters through example
> code, by running example code, you get concrete return
> values that you can analyse in the same fashion. Furthermore, Figure
> 4 (for example) shows a generated file with functions
> with return types. What am I missing here?

Some explanation provided.

> ## Typos
> 
> all over the paper: runtime vs run time vs run-time!
OK...
> 
> p. 1: "..., which is error prone and time consuming." - error-prone
> p. 2: "..., which is time consuming and error prone." - error-prone
FIXED^2
> p. 8: "A TypeScript Declaration File" - declaration file (why
> title-case?)
FIXED
> p. 8: "... so that test cases were difficult to reap from the NPM packages:..." - read
NOT FIXED
> p. 9: "The Runtime Information block in Figure 3" - run-time information (why title-case?)
FIXED
> p. 9: "For example, Function f ... Property foo ... Parameter a
> ... - function ... property ... parameter (it seems weird to start
> with upper-case letters for the itemization)
> p. 11: " and traverse the Abstract Syntax Tree of a TypeScript program" - abstract syntax tree
OK
> p. 12: "DefinitelyTyped Repository" - repository (why title-case?)
OK
> p. 12: "uses the TypeScript Compiler API" - TypeScript compiler API
> (why title-case?)
OK
> p. 18: "Excerpt of JS implementation that throws an Error when in- put parameter is not a string" - error
OK
> p. 19: "Properties of interface Result were ignored for readability"
> - result
NOT FIXED
> p. 22: "future incorporation of a Symbolic Execution Engine that refines the" - symbolic execution engine (why title-case?)
OK
> 
> ## Others
> p. 1: "from the TypeScript IDE and thus to incorrect uses of" - and, thus,  to
> p. 2: "a scripting language and thus lacks features for maintaining" - and, thus, lacks
> p. 4: "The implementation and the results are available under the following repositories:" -
> ...following repositories.
> p. 6: "It infers the overall structure of the JavaScript library,..." - non-referential _it_, why not write "the block infers..."
> p. 9: "...meaningful examples, thus avoiding possible cold start problems." - examples, thus, avoiding
ALL FIXED
> 
> 
> ----------------------- REVIEW 3 ---------------------
> SUBMISSION: 28
> TITLE: Generation of TypeScript Declaration Files from JavaScript Code
> AUTHORS: Fernando Cristiani and Peter Thiemann
> 
> ----------- Overall evaluation -----------
> SCORE: -1 (major revision)
> ----- TEXT:
> The paper presents a tool that automatically generates TypeScript
> definition files for JavaScript projects. The tool uses > dynamic
> analysis. It runs the code samples found in the README file in the
> project repository, records information about > method invocations
> and argument values and uses those to generate type definitions. The
> paper evaluates the tool by > comparing generated definitions with
> hand-written definitions from the DefinitelyTyped repository - this
> is done for libraries > which have code samples in README,
> i.e. about ~270 out of ~6000. 
> 
> The paper presents a tool that is useful, although it has a fairly
> limited scope. The approach presented by the authors - to > run a
> code sample using a library and use runtime monitoring to deduce the
> structure of the library - is definitely a promising > one. The
> authors make a good first step towards building a practical
> tool. However, the current implementation of the tool is > quite
> basic in a number of ways. First, it relies on code samples in
> README, which is not exactly common (it will work only > for quite
> basic libraries). Second, it ignores many interesting types that a
> library could use - most prominently arrays, but > those are just
> one example. 
> 
> I believe the work presented in the paper is still interesting and useful, but it would need to be much more complete to > justify writing an extensive report about it. Although the journal does not have a specific category of "Tool" papers, I imagine > that if the authors presented the work in a shorter paper and clearly acknowledged the limitations and contributions, then it > would be interesting enough for publication. (Following the idea that the length of the paper should be commensurable with > the contribution.) The current version spends a quite a lot of time discussing the high-level structure of the tool and some > practical details which I did not find particularly important.
> 
> DETAILED COMMENTS
> 
> * In the introduction, you say that manually writing definitions is error-prone, but your paper then uses manually written > definitions from DefinitelyTyped as ground truth. I find this a bit paradoxical - maybe you could instead say that this is a lot > of work and there are many libraries that are not covered by DefinitelyTyped?

ADDED a SENTENCE

> 
> * I have to admit that I was a bit confused by the idea of "templates" that is used in the paper. Do the templates determine > just how you structure the definition, or are they somehow limiting what kind of constructs can be used in a given file? I think > this would benefit from some clarification in the paper.

EXPLAINED in section 3

> * There is very little detail in the paper about how you infer types from the sample values that are observed during the > runtime execution. I guess the mechanism here is quite basic, but it is still worth discussing. (You might be interested in the > related paper [1] which does this for more complex types, but also covers things like records with optional fields.)

DONE in section 4

> * I think your evaluation is somewhat limited by the fact that the libraries that have a complete documentation in the > README are the ones with very simple API. It would be interesting to evaluate how your approach works on larger libraries > (even if that meant copying some examples into a readme manually).
> 
> * I think the phrasing that "more advanced types are overblown" is a bit unfortunate, especially if the tool does not support > some quite basic types like arrays.

TONED DOWN. We have arrays now.

> * The paper should give a much more extensive overview of related work - for example, it should discuss what has been > done using static analyses, whether similar tools exist for other languages and there might also be other interesting related > approaches (perhaps some based on machine learning methods). The references in the submission are fairly minimal.
> 
> [1] T. Petricek, G. Guera, D. Syme. Types from data: making structured data first-class citizens in F#, PLDI 2016